---
title: "Council_Seasonal_Annual_Flux_Budgets_Met_Means"
output: html_document
date: "2025-12-17"
#Important notes:
  # - includes SWC_3 when TS_3_gf (gapfilled soil temp) is > 2 to avoid frozen ground / freezing processes 
  # -includes winters that span across 2 years (winter of 2021 inclues winter 2021 - end of winter in 2022)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)


Sys.setenv(TZ='UTC')
```

# Set working directory and then Load data

```{r}

#HH daataset
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

#needs continuous day variable - daily avg, calculated in processing steps - using ".4" with GF temps, two WD classifications, and corrected seasonal delineations 
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.4.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

```


#Set year colors for consistency 
```{r}
# Define consistent year colors
year_colors <- c("2017" = "#E41A1C",  # Red
                 "2018" = "#377EB8",  # Blue
                 "2019" = "#4DAF4A",  # Green
                 "2020" = "#984EA3",  # Purple
                 "2021" = "#FF7F00",  # Orange
                 "2022" = "#A65628")  # Brown

# Convert year to factor for consistent color mapping
df_avg$year_factor <- as.factor(df_avg$year)
```

#SWC when ground is not frozen - SWC measurements in frozen ground are not reliable with specific calibrations or data processing 

####Create variable for SWC_3 means when winter is excluded and TS_3_gf is > 2 degC = SWC_3_nf
```{r}
df_avg <- df_avg %>%
  mutate(
    # SWC_3_nf: Only use SWC when soil is not frozen (TS > 2°C)
    # Set to NA when TS <= 2 or when it's winter season
    SWC_3_nf = case_when(
      season == "Winter" ~ NA_real_,              # Exclude all winter
      TS_3_gapfilled <= 2 ~ NA_real_,             # Exclude frozen soil
      TRUE ~ SWC_3_1_1                            # Use actual SWC otherwise
    )
  )

#Data inclusion/exclusion diagnostics: 
cat(sprintf("  Original SWC observations: %d\n", sum(!is.na(df_avg$SWC_3_1_1))))
cat(sprintf("  Non-frozen SWC observations: %d\n", sum(!is.na(df_avg$SWC_3_nf))))
cat(sprintf("  Excluded: %d observations\n\n", 
            sum(!is.na(df_avg$SWC_3_1_1)) - sum(!is.na(df_avg$SWC_3_nf))))

#check
ggplot(df_avg %>%
         filter(season != "Winter" & TS_3_gapfilled > 2)) +
  geom_point(aes(x = TS_3_gapfilled, y = SWC_3_1_1)) +
  geom_point(aes(TS_3_gapfilled, y = SWC_3_nf), color = "red", shape = 5)

#check
ggplot(df_avg) +
  geom_point(aes(x = TS_3_gapfilled, y = SWC_3_1_1)) +
  geom_point(aes(TS_3_gapfilled, y = SWC_3_nf), color = "red", shape = 5)


```

#Table 1 in Council paper: seasonal and annual C fluxes, total C budget & met variables

#Annual Budgets / Means 

#Annual - not winter-adj
```{r}
library(dplyr)
library(lubridate)

#Annual and seasonal summaries with SE 

#no SE for CO2 and CH4 fluxes as these are sums **** then we calc averages with SE for the met variables 

annual_budgets <- df_avg %>%
  group_by(year = year(date)) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete year data 
  summarize(
    #pre winter-adj
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    #pre winter-adj 
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    #pre winter-adj
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    # GPP budget
    GPP_budget = sum(GPP_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    # pre-winter adj RECO budget
    RECO_budget = sum(RECO_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    #measured, not GF
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    #SWC_3 all 
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    #SWC_3_nf (not frozen) 
    mean_soil_moisture_nf = mean(SWC_3_nf, na.rm = TRUE),
    se_soil_moisture_nf = sd(SWC_3_nf, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_nf))),
    #measured TA, not GF 
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    #ERA-mod GF TA
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
    se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
   #ERA-mod GF TS_3
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled)))),
   #vapor pressure deficit
    mean_VPD = mean(VPD, na.rm = TRUE),
    se_VPD = sd(VPD, na.rm = TRUE) / sqrt(sum(!is.na(VPD))),
   #relative humidity 
    mean_RH = mean(RH, na.rm = TRUE),
    se_RH = sd(RH, na.rm = TRUE) / sqrt(sum(!is.na(RH))),
    # Sensible heat
    mean_H = mean(H, na.rm = TRUE),
    se_H = sd(H, na.rm = TRUE) / sqrt(sum(!is.na(H))),
    # Latent heat
    mean_LE = mean(LE, na.rm = TRUE),
    se_LE = sd(LE, na.rm = TRUE) / sqrt(sum(!is.na(LE))),
    # Ground heat
    mean_G = mean(G_1_1_1, na.rm = TRUE),
    se_G = sd(G_1_1_1, na.rm = TRUE) / sqrt(sum(!is.na(G_1_1_1))),
   
    days = n()
      )

print(annual_budgets)
```

#Create Table 
```{r}
library(tidyverse)

#Create formatted columns with mean ± SE

annual_budgets_formatted <- annual_budgets %>%
  arrange(year) %>%  # Sort by yr
  mutate(
    # Carbon budgets (no SE, just values)
    `CO2 Budget` = sprintf("%.2f", CO2_budget),
    `CH4 Budget` = sprintf("%.2f", CH4_budget),
    `Total C Budget` = sprintf("%.2f", total_C_budget),
    `GPP Budget` = sprintf("%.2f", GPP_budget),
    `RECO Budget` = sprintf("%.2f", RECO_budget),
    
    # Temperature variables
    `Soil Temp` = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    `Air Temp` = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    `Air Temp (GF)` = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    `Soil Temp (GF)` = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
    
    # Soil moisture
    `SWC` = sprintf("%.2f ± %.2f", mean_soil_moisture, se_soil_moisture),
    `SWC (non-frozen)` = sprintf("%.2f ± %.2f", mean_soil_moisture_nf, se_soil_moisture_nf),
    
    # Other met variables
    `VPD` = sprintf("%.2f ± %.2f", mean_VPD, se_VPD),
    `RH` = sprintf("%.2f ± %.2f", mean_RH, se_RH),
    `H` = sprintf("%.2f ± %.2f", mean_H, se_H),
    `LE` = sprintf("%.2f ± %.2f", mean_LE, se_LE),
    `G` = sprintf("%.2f ± %.2f", mean_G, se_G),
    
    # Keep identifiers and sample size
    Year = year,
    Days = days
  ) %>%
  # Select formatted columns in desired order
  select(Year, Days, 
         `CO2 Budget`, `CH4 Budget`, `Total C Budget`, `GPP Budget`, `RECO Budget`,
         `Air Temp (GF)`, `Soil Temp (GF)`,
         `SWC (non-frozen)`,
         `VPD`, `RH`, `H`, `LE`, `G`)

# Display as simple table
print(annual_budgets_formatted)


```


#Save table 
```{r}
# # Save to Excel
 library(writexl)
write_xlsx(annual_budgets_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/annual_means_budgets_12.17.2025.xlsx")
```




#Seasonal by year - not winter adj - add in SWC_3_nf 
```{r}
# Seasonal summaries with SE &  with total C budget

seasonal_budgets <- df_avg %>%
  group_by(year = year(date), season) %>%
  #filter(year != 2023) %>% #keeping 2023 winter
  summarize(
    #daily avg NEE flux (FC_F)
   mean_daily_NEE = mean(FC_F, na.rm = TRUE),
    se_NEE = sd(FC_F, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
   #daily avg FCH4_F flux
    mean_daily_FCH4 = mean(FCH4_F, na.rm = TRUE),
    se_FCH4 = sd(FCH4_F, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
   # pre-winter adj CO2 budget
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
   #pre-winter adj CH4 budget 
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
   #pre-winter adj total C budget
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    # GPP budget
    GPP_budget = sum(GPP_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
   #daily avg GPP flux (GPP_F)
    mean_daily_GPP = mean(GPP_F, na.rm = TRUE),
    se_GPP = sd(GPP_F, na.rm = TRUE)/sqrt(sum(!is.na(GPP_F))),
   # pre-winter adj RECO budget
    RECO_budget = sum(RECO_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    #daily avg RECO flux (RECO_F)
    mean_daily_RECO = mean(RECO_F, na.rm = TRUE),
    se_RECO = sd(RECO_F, na.rm = TRUE)/sqrt(sum(!is.na(RECO_F))),
      #measured soil temp, no GF
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
   #SWC_3, all 
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    #SWC_3, excluding winter and TS_3_gf > 2 degC
    mean_soil_moisture_nf = mean(SWC_3_nf, na.rm = TRUE),
    se_soil_moisture_nf = sd(SWC_3_nf, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_nf))),
   #measured TA, no GF
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
   #ERA-mod GF TA
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
    se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
   #ERA-mod GF TS_3
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled)))),
   #vapor pressure deficit
    mean_VPD = mean(VPD, na.rm = TRUE),
    se_VPD = sd(VPD, na.rm = TRUE) / sqrt(sum(!is.na(VPD))),
   #relative humidity 
    mean_RH = mean(RH, na.rm = TRUE),
    se_RH = sd(RH, na.rm = TRUE) / sqrt(sum(!is.na(RH))),
    # Sensible heat
    mean_H = mean(H, na.rm = TRUE),
    se_H = sd(H, na.rm = TRUE) / sqrt(sum(!is.na(H))),
    # Latent heat
    mean_LE = mean(LE, na.rm = TRUE),
    se_LE = sd(LE, na.rm = TRUE) / sqrt(sum(!is.na(LE))),
    # Ground heat
    mean_G = mean(G_1_1_1, na.rm = TRUE),
    se_G = sd(G_1_1_1, na.rm = TRUE) / sqrt(sum(!is.na(G_1_1_1))),
    days = n()
  )


print(seasonal_budgets)
```
#Create Table - not winter adj 
```{r}
library(tidyverse)


#Create formatted columns with mean ± SE

seasonal_budgets_formatted <- seasonal_budgets %>%
  mutate(
    season = factor(season,
                   levels = c("Growing Season", "Fall Senescence", "Winter"))
  ) %>%
  arrange(year, season) %>%  # Sort by this order
  mutate(
    # Fluxes with ± SE
    `Daily NEE` = sprintf("%.2f ± %.2f", mean_daily_NEE, se_NEE),
    `Daily FCH4` = sprintf("%.2f ± %.2f", mean_daily_FCH4, se_FCH4),
     `Daily GPP` = sprintf("%.2f ± %.2f", mean_daily_GPP, se_GPP),
     `Daily RECO` = sprintf("%.2f ± %.2f", mean_daily_RECO, se_RECO),
    
    # Carbon budgets (no SE, just values)
    `CO2 Budget` = sprintf("%.2f", CO2_budget),
    `CH4 Budget` = sprintf("%.2f", CH4_budget),
    `Total C Budget` = sprintf("%.2f", total_C_budget),
     `GPP Budget` = sprintf("%.2f", GPP_budget),
     `RECO Budget` = sprintf("%.2f", RECO_budget),
    
    # Temperature variables
    `Soil Temp` = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    `Air Temp` = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    `Air Temp (GF)` = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    `Soil Temp (GF)` = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
    
    # Soil moisture
    `SWC` = sprintf("%.3f ± %.3f", mean_soil_moisture, se_soil_moisture),
    `SWC (non-frozen)` = sprintf("%.3f ± %.3f", mean_soil_moisture_nf, se_soil_moisture_nf),
    
    # Other met variables
    `VPD` = sprintf("%.2f ± %.2f", mean_VPD, se_VPD),
    `RH` = sprintf("%.2f ± %.2f", mean_RH, se_RH),
    `H` = sprintf("%.2f ± %.2f", mean_H, se_H),
    `LE` = sprintf("%.2f ± %.2f", mean_LE, se_LE),
    `G` = sprintf("%.2f ± %.2f", mean_G, se_G),
    
    # Keep identifiers and sample size
    Year = year,
    Season = season,
    Days = days
  ) %>%
  # Select formatted columns in desired order
  select(Year, Season, Days, 
         `Daily NEE`, `Daily FCH4`, `Daily GPP`, `Daily RECO`,
         `CO2 Budget`, `CH4 Budget`, `Total C Budget`, `GPP Budget`, `RECO Budget`,
         `Air Temp (GF)`, `Soil Temp (GF)`,
         `SWC (non-frozen)`,
         `VPD`, `RH`, `H`, `LE`, `G`)

# Display as simple table
print(seasonal_budgets_formatted)




```
#Save table 
```{r}
# # Save to Excel
 library(writexl)
write_xlsx(seasonal_budgets_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/seasonal_means_budgets_12.17.2025.xlsx")
```



#Create Winter-year for cross-year winters
```{r}
#Use July as a cutoff. Winter_year_2021 uses all DOY assigned "Winter" AFTER July 2021 and all DOY assigned "Winter" in 2022 BEFORE month of July 
#example: Winter 2021 = Winter obs from July-Dec 2021 + Winter obs from Jan-June 2022

df_avg2 <- df_avg %>%
  mutate(
    year = year(date),
    month = month(date),
    doy = yday(date)
  )



# Assign winter_year based on season and July cutoff


df_avg2 <- df_avg2 %>%
  mutate(
    winter_year = case_when(
      # For NON-WINTER seasons: use calendar year
      season != "Winter" ~ year,
      
      # For WINTER season AFTER July (months 7-12): THIS year's winter
      season == "Winter" & month >= 7 ~ year,
      
      # For WINTER season BEFORE July (months 1-6): PREVIOUS year's winter
      season == "Winter" & month < 7 ~ year - 1,
      
      # Catch-all
      TRUE ~ year
    )
  )

#Verify
winter_verification <- df_avg2 %>%
  filter(season == "Winter") %>%
  group_by(winter_year) %>%
  summarise(
    n_obs = n(),
    calendar_years = paste(sort(unique(year)), collapse = " & "),
    months_included = paste(sort(unique(month)), collapse = ", "),
    start_date = min(date),
    end_date = max(date),
    date_range_desc = paste(format(min(date), "%b %d, %Y"), "to", 
                           format(max(date), "%b %d, %Y")),
    .groups = "drop"
  ) %>%
  arrange(winter_year)

print(winter_verification)



```
#Winter budgets and means based on winter year
```{r}

winter_year_budgets <- df_avg2 %>%
  filter(season == "Winter") %>%
  group_by(winter_year) %>%
  summarize(
    # Fluxes
    mean_daily_NEE = mean(FC_F, na.rm = TRUE),
    se_NEE = sd(FC_F, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
    mean_daily_FCH4 = mean(FCH4_F, na.rm = TRUE),
    se_FCH4 = sd(FCH4_F, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
    
    # Carbon budgets
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    total_C_budget = CO2_budget + CH4_budget,
    
    # Temperature
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_gapfilled = mean(TA_gapfilled, na.rm = TRUE),
    se_TA_gapfilled = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))),
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE),
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_gapfilled))),
    
    # Soil moisture - both versions
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_SWC_nf = mean(SWC_3_nf, na.rm = TRUE),
    se_SWC_nf = sd(SWC_3_nf, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_nf))),
    n_SWC_nf = sum(!is.na(SWC_3_nf)),
    
    # Other met variables
    mean_VPD = mean(VPD, na.rm = TRUE),
    se_VPD = sd(VPD, na.rm = TRUE) / sqrt(sum(!is.na(VPD))),
    mean_RH = mean(RH, na.rm = TRUE),
    se_RH = sd(RH, na.rm = TRUE) / sqrt(sum(!is.na(RH))),
    mean_H = mean(H, na.rm = TRUE),
    se_H = sd(H, na.rm = TRUE) / sqrt(sum(!is.na(H))),
    mean_LE = mean(LE, na.rm = TRUE),
    se_LE = sd(LE, na.rm = TRUE) / sqrt(sum(!is.na(LE))),
    mean_G = mean(G_1_1_1, na.rm = TRUE),
    se_G = sd(G_1_1_1, na.rm = TRUE) / sqrt(sum(!is.na(G_1_1_1))),
    
    days = n(),
    .groups = "drop"
  ) %>%
  rename(year = winter_year)  # Rename for consistency

print(winter_year_budgets)


#checking 
winter_year_2019 <- df_avg2 %>%
  filter(winter_year == "2019" & season == "Winter") #ok looks good 

winter_year_2019_budgets <- winter_year_2019 %>%
  summarize(
    # Fluxes
    mean_daily_NEE = mean(FC_F, na.rm = TRUE),
    se_NEE = sd(FC_F, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
    mean_daily_FCH4 = mean(FCH4_F, na.rm = TRUE),
    se_FCH4 = sd(FCH4_F, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
    
    # Carbon budgets
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
 
          days = n(),
    .groups = "drop"
  )
```

#**** Note on difference:
####FC_F remains similar between the old winter calcs and winter_year calcs, but FCH4_F *doubled* - not sure why FCH4_F would change so much but FC_F wouldn't, so running visual checks 
```{r}

#checking 
winter_year_2019_FCH4 <- winter_year_2019 %>%
  filter(winter_year == "2019" & season == "Winter") #ok looks good 

winter_year_2019_budgets <- winter_year_2019 %>%
  summarize(
    # Fluxes
    mean_daily_NEE = mean(FC_F, na.rm = TRUE),
    se_NEE = sd(FC_F, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
    mean_daily_FCH4 = mean(FCH4_F, na.rm = TRUE),
    se_FCH4 = sd(FCH4_F, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
    
    # Carbon budgets
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
 
          days = n(),
    .groups = "drop"
  )

print(winter_year_budgets)
print(winter_year_2019_budgets)

#Visual check of difference between 2019 winter and 2019 winter_year
#FCH4_F
ggplot(data = winter_year_2019) +
  geom_point(aes(x = date, y = FCH4_F, color = year_factor))

ggplot(data = df_avg2 %>%
         filter(year == "2019" & season == "Winter")) +
    geom_point(aes(x = date, y = FCH4_F, color = year))

#FC_F
ggplot(data = winter_year_2019) +
  geom_point(aes(x = date, y = FC_F, color = year_factor)) +
  labs(title = "2019 Winter year")

ggplot(data = df_avg2 %>%
         filter(year == "2019" & season == "Winter")) +
    geom_point(aes(x = date, y = FC_F, color = year)) +
  labs(title = "df_avg_2019_winter")

df_avg_2019_winter <- df_avg2 %>%
  filter(year == "2019" & season == "Winter")

summary(df_avg_2019_winter$FCH4_F)
summary(winter_year_2019$FCH4_F)

summary(df_avg_2019_winter$FC_F)
summary(winter_year_2019$FC_F)
```



#Code for calculating budgets based on all seasons & assigned winter year 
```{r}
seasonal_budgets_winteryear <- df_avg2 %>%
  group_by(winter_year, season) %>%
  summarize(
    #daily avg NEE flux (FC_F)
   mean_daily_NEE = mean(FC_F, na.rm = TRUE),
    se_NEE = sd(FC_F, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
   #daily avg FCH4_F flux
    mean_daily_FCH4 = mean(FCH4_F, na.rm = TRUE),
    se_FCH4 = sd(FCH4_F, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
   # pre-winter adj CO2 budget
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
   #pre-winter adj CH4 budget 
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
   #pre-winter adj total C budget
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    # GPP budget
    GPP_budget = sum(GPP_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
   #daily avg GPP flux (GPP_F)
    mean_daily_GPP = mean(GPP_F, na.rm = TRUE),
    se_GPP = sd(GPP_F, na.rm = TRUE)/sqrt(sum(!is.na(GPP_F))),
   # pre-winter adj RECO budget
    RECO_budget = sum(RECO_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    #daily avg RECO flux (RECO_F)
    mean_daily_RECO = mean(RECO_F, na.rm = TRUE),
    se_RECO = sd(RECO_F, na.rm = TRUE)/sqrt(sum(!is.na(RECO_F))),
      #measured soil temp, no GF
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
   #SWC_3, all 
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    #SWC_3, excluding winter and TS_3_gf > 2 degC
    mean_soil_moisture_nf = mean(SWC_3_nf, na.rm = TRUE),
    se_soil_moisture_nf = sd(SWC_3_nf, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_nf))),
   #measured TA, no GF
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
   #ERA-mod GF TA
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
    se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
   #ERA-mod GF TS_3
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled)))),
   #vapor pressure deficit
    mean_VPD = mean(VPD, na.rm = TRUE),
    se_VPD = sd(VPD, na.rm = TRUE) / sqrt(sum(!is.na(VPD))),
   #relative humidity 
    mean_RH = mean(RH, na.rm = TRUE),
    se_RH = sd(RH, na.rm = TRUE) / sqrt(sum(!is.na(RH))),
    # Sensible heat
    mean_H = mean(H, na.rm = TRUE),
    se_H = sd(H, na.rm = TRUE) / sqrt(sum(!is.na(H))),
    # Latent heat
    mean_LE = mean(LE, na.rm = TRUE),
    se_LE = sd(LE, na.rm = TRUE) / sqrt(sum(!is.na(LE))),
    # Ground heat
    mean_G = mean(G_1_1_1, na.rm = TRUE),
    se_G = sd(G_1_1_1, na.rm = TRUE) / sqrt(sum(!is.na(G_1_1_1))),
  
     # Winter year Date info
    wy_start_date = min(date),
    wy_end_date = max(date),
    
   #days per season 
   days = n(),
    .groups = "drop"
  ) %>%
  rename(year = winter_year) 

print(seasonal_budgets_winteryear)
```

#Make table with winter-year budgets - not winter-adj

```{r}
library(tidyverse)


#Create formatted columns with mean ± SE

seasonal_budgets_winteryear_formatted <- seasonal_budgets_winteryear %>%
  mutate(
    season = factor(season,
                   levels = c("Growing Season", "Fall Senescence", "Winter"))
  ) %>%
  arrange(year, season) %>%  # Sort by this order
  mutate(
    # Fluxes with ± SE
    `Daily NEE` = sprintf("%.2f ± %.2f", mean_daily_NEE, se_NEE),
    `Daily FCH4` = sprintf("%.2f ± %.2f", mean_daily_FCH4, se_FCH4),
     `Daily GPP` = sprintf("%.2f ± %.2f", mean_daily_GPP, se_GPP),
     `Daily RECO` = sprintf("%.2f ± %.2f", mean_daily_RECO, se_RECO),
    
    # Carbon budgets (no SE, just values)
    `CO2 Budget` = sprintf("%.2f", CO2_budget),
    `CH4 Budget` = sprintf("%.2f", CH4_budget),
    `Total C Budget` = sprintf("%.2f", total_C_budget),
     `GPP Budget` = sprintf("%.2f", GPP_budget),
     `RECO Budget` = sprintf("%.2f", RECO_budget),
    
    # Temperature variables
    `Soil Temp` = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    `Air Temp` = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    `Air Temp (GF)` = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    `Soil Temp (GF)` = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
    
    # Soil moisture
    `SWC` = sprintf("%.3f ± %.3f", mean_soil_moisture, se_soil_moisture),
    `SWC (non-frozen)` = sprintf("%.3f ± %.3f", mean_soil_moisture_nf, se_soil_moisture_nf),
    
    # Other met variables
    `VPD` = sprintf("%.2f ± %.2f", mean_VPD, se_VPD),
    `RH` = sprintf("%.2f ± %.2f", mean_RH, se_RH),
    `H` = sprintf("%.2f ± %.2f", mean_H, se_H),
    `LE` = sprintf("%.2f ± %.2f", mean_LE, se_LE),
    `G` = sprintf("%.2f ± %.2f", mean_G, se_G),
    
    # Keep identifiers and sample size
    Year = year,
    Season = season,
    Days = days
  ) %>%
  # Select formatted columns in desired order
  select(Year, Season, Days, 
         `Daily NEE`, `Daily FCH4`, `Daily GPP`, `Daily RECO`,
         `CO2 Budget`, `CH4 Budget`, `Total C Budget`, `GPP Budget`, `RECO Budget`,
         `Air Temp (GF)`, `Soil Temp (GF)`,
         `SWC (non-frozen)`,
         `VPD`, `RH`, `H`, `LE`, `G`)

# Display as simple table
print(seasonal_budgets_winteryear_formatted)



```


#Save table 
```{r}
# # Save to Excel
 library(writexl)
write_xlsx(seasonal_budgets_winteryear_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/seasonal_means_budgets_winteryear_12.17.2025.xlsx")
```

# % gapfilled - edit 
#### comparing HH and df_avg %gapfilled differ by only ~1%, so using df_avg for calculating gf%



```{r}

gapfill_summary <- df_avg2 %>%
  # Just group directly by winter_year (which already exists!)
  group_by(winter_year, season) %>%
  summarise(
    # Total days in this season
    total_days = n(),
    
    # NEE (FC vs FC_F)
    NEE_measured = sum(!is.na(FC)),
    NEE_gapfilled = sum(is.na(FC) & !is.na(FC_F)),
    NEE_pct_gapfilled = round((NEE_gapfilled / total_days) * 100, 1),
    
    # FCH4 (FCH4 vs FCH4_F)
    FCH4_measured = sum(!is.na(FCH4)),
    FCH4_gapfilled = sum(is.na(FCH4) & !is.na(FCH4_F)),
    FCH4_pct_gapfilled = round((FCH4_gapfilled / total_days) * 100, 1),
    
    # GPP (GPP vs GPP_F)
    GPP_measured = sum(!is.na(GPP)),
    GPP_gapfilled = sum(is.na(GPP) & !is.na(GPP_F)),
    GPP_pct_gapfilled = round((GPP_gapfilled / total_days) * 100, 1),
    
    # RECO (RECO vs RECO_F)
    RECO_measured = sum(!is.na(RECO)),
    RECO_gapfilled = sum(is.na(RECO) & !is.na(RECO_F)),
    RECO_pct_gapfilled = round((RECO_gapfilled / total_days) * 100, 1),
    
    # Air Temperature (TA vs TA_gapfilled)
    TA_measured = sum(!is.na(TA)),
    TA_gapfilled = sum(is.na(TA) & !is.na(TA_gapfilled)),
    TA_pct_gapfilled = round((TA_gapfilled / total_days) * 100, 1),
    
    # Soil Temperature (TS_3_1_1 vs TS_3_gapfilled)
    TS_measured = sum(!is.na(TS_3_1_1)),
    TS_gapfilled = sum(is.na(TS_3_1_1) & !is.na(TS_3_gapfilled)),
    TS_pct_gapfilled = round((TS_gapfilled / total_days) * 100, 1),
    
    .groups = "drop"
  ) %>%
  # Rename to "year" for consistency with other tables
  rename(year = winter_year)



# Display summary
print(gapfill_summary)


#compact table with only %gf

gapfill_pct_only <- gapfill_summary %>%
  select(year, season, total_days,
         NEE_pct_gapfilled,
         FCH4_pct_gapfilled,
         GPP_pct_gapfilled,
         RECO_pct_gapfilled,
         TA_pct_gapfilled,
         TS_pct_gapfilled)

print(gapfill_pct_only)

#visual check - FC
ggplot(data = df_avg_2019_winter) +
  geom_point(aes(x = date, y = FC)) +
  geom_point(aes(x = date, y = FC_F, color = "red")) +
  labs(title = "df_avg_2019_winter")

ggplot(data = winter_year_2019) +
  geom_point(aes(x = date, y = FC)) +
  geom_point(aes(x = date, y = FC_F, color = "red"))+
  labs(title = "2019 winter year")

#visual check with temp
ggplot(data = df_avg_2019_winter) +
  geom_point(aes(x = date, y = TA)) +
  geom_jitter(aes(x = date, y = TA_gapfilled, color = "red")) +
  labs(title = "df_avg_2019_winter")

sum(is.na(df_avg_2019_winter$TA))

ggplot(data = winter_year_2019) +
  geom_point(aes(x = date, y = TA)) +
  geom_jitter(aes(x = date, y = TA_gapfilled, color = "red"))+
  labs(title = "2019 winter year")

sum(is.na(winter_year_2019$TA))

#yup, checks out 


ggplot(data = df_avg %>%
         filter(year == "2018" & season == "Winter")) +
  geom_point(aes(x = date, y = FC)) +
  geom_jitter(aes(x = date, y = FC_F, color = "red"))+
  labs(title = "2018 winter year")

winter_year_2018 <- df_avg2 %>%
  filter(winter_year == "2018" & season == "Winter")

winter_year_2022 <- df_avg2 %>%
  filter(winter_year == "2022" & season == "Winter")

ggplot(data = df_avg2 %>%
         filter(winter_year == "2018" & season == "Winter")) +
  geom_point(aes(x = date, y = FC)) +
  geom_jitter(aes(x = date, y = FC_F, color = "red"))+
  labs(title = "df_avg_2018 winter")


sum(is.na(winter_year_2018$FC))
sum(is.na(winter_year_2019$FC))
sum(is.na(winter_year_2022$FCH4))

```
#merge %gf with seasonal summaries by winter_year
```{r}

seasonal_budgets_winteryear_with_gf <- seasonal_budgets_winteryear %>%
  left_join(
    gapfill_pct_only,
    by = c("year", "season")
  )


# Display merged data
print(seasonal_budgets_winteryear_with_gf %>%
      select(year, season, days, 
             mean_daily_NEE, NEE_pct_gapfilled,
             mean_daily_FCH4, FCH4_pct_gapfilled,
             ))

#formatted table 

seasonal_table_with_gf <- seasonal_budgets_winteryear_with_gf %>%
  mutate(
    # Format mean ± SE with % gapfilled
    `NEE` = sprintf("%.2f ± %.2f (%s%%)", 
                    mean_daily_NEE, se_NEE, NEE_pct_gapfilled),
    `FCH4` = sprintf("%.2f ± %.2f (%s%%)", 
                     mean_daily_FCH4, se_FCH4, FCH4_pct_gapfilled),
    `GPP` = sprintf("%.2f ± %.2f (%s%%)",
                    mean_daily_GPP, se_GPP, GPP_pct_gapfilled),
    `RECO` = sprintf("%.2f ± %.2f (%s%%)",
                     mean_daily_RECO, se_RECO, RECO_pct_gapfilled),
    `Air Temp` = sprintf("%.2f ± %.2f (%s%%)",
                         mean_TA_gapfilled, se_TA_gapfilled, TA_pct_gapfilled),
    `Soil Temp` = sprintf("%.2f ± %.2f (%s%%)",
                          mean_TS_3_gapfilled, se_TS_3_gapfilled, TS_pct_gapfilled),
    
    Year = year,
    Season = season,
    Days = days
  ) %>%
  select(Year, Season, Days, NEE, FCH4, GPP, RECO, `Air Temp`, `Soil Temp`)

seasonal_table_with_gf 

```
#save table with %gf
```{r}

 library(writexl)
write_xlsx(seasonal_table_with_gf , "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/seasonal_means_budgets_winteryear_percentGF_12.17.2025.xlsx")

```



#*** NEW WINTER ADJ BASED ON WINTER YEAR

#Winter adj based on proportion of 2018 daily mean winter flux : daily mean flux of growing season 

####FCH4 - daily mean winter adj 

#based on 2018 daily avg growing season fluxes, winter daily avg CH4 flux can be ~44% of growing season daily avg flux (0.443) --> multiply winter daily avg fluxes by 0.443 & re-calc the daily avg CH4 flux to get an adjusted potential winter daily avg flux 



```{r}

library(tidyverse)

# 1. Calculate growing season means AND SE for each winter_year

growing_season_means <- df_avg2 %>%
  filter(season == "Growing Season") %>%
  group_by(winter_year) %>%
  summarise(
    # Means
    mean_growing_NEE = mean(FC_F, na.rm = TRUE),
    mean_growing_FCH4 = mean(FCH4_F, na.rm = TRUE),
    
    # Standard Errors
    se_growing_NEE = sd(FC_F, na.rm = TRUE) / sqrt(sum(!is.na(FC_F))),
    se_growing_FCH4 = sd(FCH4_F, na.rm = TRUE) / sqrt(sum(!is.na(FCH4_F))),
    
    # Sample size
    n_growing = n(),
    
    .groups = "drop"
  )


print(growing_season_means %>%
      mutate(
        `NEE ± SE` = sprintf("%.3f ± %.3f", mean_growing_NEE, se_growing_NEE),
        `FCH4 ± SE` = sprintf("%.4f ± %.4f", mean_growing_FCH4, se_growing_FCH4)
      ) %>%
      select(winter_year, `NEE ± SE`, `FCH4 ± SE`, n_growing))


# 2. Count winter days for each winter_year

winter_days <- df_avg2 %>%
  filter(season == "Winter") %>%
  group_by(winter_year) %>%
  summarise(
    winter_days = n(),
    .groups = "drop"
  )

print(winter_days)


# 3: Calculate winter-adjusted budgets WITH PROPAGATED SE


# Define conversion factors
CO2_conversion <- 60 * 60 * 24 * (1/1000000) * 12  # = 1.0368; umol/m2/s to g/m2/day
CH4_conversion <- 60 * 60 * 24 * (1/1000000000) * 12  # = 0.0010368; nmol/m2/s to g/m2/day

# Define proportion factors - based on year 2018 with best winter coverage 
NEE_winter_proportion <- 0.451
FCH4_winter_proportion <- 0.443


# Calculate winter-adjusted budgets with SE propagation
winter_adj_budgets <- growing_season_means %>%
  left_join(winter_days, by = "winter_year") %>%
  mutate(
    # ========================================
    # Winter-adjusted DAILY flux (μmol/m2/s or nmol/m²/s)
    # ========================================
    
    # Mean daily flux (still in original units)
    winter_adj_daily_NEE_flux = abs(mean_growing_NEE) * NEE_winter_proportion,
    winter_adj_daily_FCH4_flux = abs(mean_growing_FCH4) * FCH4_winter_proportion,
    
    # SE for daily flux: SE_new = SE_original × proportion
    # (When multiplying by a constant, SE scales by that constant)
    se_winter_adj_daily_NEE_flux = se_growing_NEE * NEE_winter_proportion,
    se_winter_adj_daily_FCH4_flux = se_growing_FCH4 * FCH4_winter_proportion,
    
    # ========================================
    # Convert daily flux to daily BUDGET (g C/m2/day)
    # ========================================
    
    # Mean daily budget
    winter_adj_daily_NEE_budget = winter_adj_daily_NEE_flux * CO2_conversion,
    winter_adj_daily_FCH4_budget = winter_adj_daily_FCH4_flux * CH4_conversion,
    
    # SE for daily budget: SE_new = SE_flux × conversion_factor
    se_winter_adj_daily_NEE_budget = se_winter_adj_daily_NEE_flux * CO2_conversion,
    se_winter_adj_daily_FCH4_budget = se_winter_adj_daily_FCH4_flux * CH4_conversion,
    
    # ========================================
    # Seasonal BUDGET (g C/m2) for entire winter)
    # ========================================
    
    # winter budget 
    winter_adj_CO2_budget = winter_adj_daily_NEE_budget * winter_days,
    winter_adj_CH4_budget = winter_adj_daily_FCH4_budget * winter_days,
    
    # SE for seasonal budget: SE_new = SE_daily × sqrt(n_days) but also saw another forum that said to just multiply by the same constant number as multiplying the mean ** need to look up 
    # This accounts for the fact that we're summing n independent days
     # se_winter_adj_CO2_budget = se_winter_adj_daily_NEE_budget * sqrt(winter_days),
     # se_winter_adj_CH4_budget = se_winter_adj_daily_FCH4_budget * sqrt(winter_days),
    
    #using just winter days as a constant is more appropriate here bc we are treating the daily avg as the overall avg and are scaling up, not treating each day as indepedent 
       se_winter_adj_CO2_budget = se_winter_adj_daily_NEE_budget * winter_days,
      se_winter_adj_CH4_budget = se_winter_adj_daily_FCH4_budget *  winter_days,
    
    # Total C budget
    winter_adj_total_C_budget = winter_adj_CO2_budget + winter_adj_CH4_budget,
    
    # SE for total C: SE_total = sqrt(SE_CO2² + SE_CH4²)
    # (When adding two variables, variances add, so SE = sqrt(SE1² + SE2²))
     se_winter_adj_total_C_budget = sqrt(se_winter_adj_CO2_budget^2 + 
                                         se_winter_adj_CH4_budget^2)
  ) %>%
  rename(year = winter_year)



#daily fluxes with SE 

winter_adj_budgets %>%
  mutate(
    `NEE (μmol m⁻² s⁻¹)` = sprintf("%.3f ± %.3f", 
                                     winter_adj_daily_NEE_flux, 
                                     se_winter_adj_daily_NEE_flux),
    `FCH4 (nmol m⁻² s⁻¹)` = sprintf("%.4f ± %.4f",
                                      winter_adj_daily_FCH4_flux,
                                      se_winter_adj_daily_FCH4_flux)
  ) %>%
  select(year, `NEE (μmol m⁻² s⁻¹)`, `FCH4 (nmol m⁻² s⁻¹)`) %>%
  print()


# Display seasonal budgets with SE
 winter_adj_budgets %>%
  mutate(
    `CO2 Budget (g C m⁻²)` = sprintf("%.2f ± %.2f",
                                      winter_adj_CO2_budget,
                                      se_winter_adj_CO2_budget),
    `CH4 Budget (g C m⁻²)` = sprintf("%.3f ± %.3f",
                                      winter_adj_CH4_budget,
                                      se_winter_adj_CH4_budget),
    `Total C Budget (g C m⁻²)` = sprintf("%.2f ± %.2f",
                                          winter_adj_total_C_budget,
                                          se_winter_adj_total_C_budget)
  ) %>%
  select(year, winter_days, `NEE (μmol m⁻² s⁻¹)`, `FCH4 (nmolm⁻² s⁻¹)`,
         `CO2 Budget (g C m⁻²)`, `CH4 Budget (g C m⁻²)`, 
         `Total C Budget (g C m⁻²)`) %>%
  print()


#Combined table 
 # Combined table with all info
combined_table <- winter_adj_budgets %>%
  mutate(
    `NEE (μmol m⁻² s⁻¹)` = sprintf("%.2f ± %.2f", 
                                     winter_adj_daily_NEE_flux, 
                                     se_winter_adj_daily_NEE_flux),
    `FCH4 (nmol m⁻² s⁻¹)` = sprintf("%.2f ± %.2f",
                                      winter_adj_daily_FCH4_flux,
                                      se_winter_adj_daily_FCH4_flux),
    `CO2 Budget (g C m⁻²)` = sprintf("%.2f ± %.2f",
                                      winter_adj_CO2_budget,
                                      se_winter_adj_CO2_budget),
    `CH4 Budget (g C m⁻²)` = sprintf("%.2f ± %.2f",
                                      winter_adj_CH4_budget,
                                      se_winter_adj_CH4_budget),
    `Total C Budget (g C m⁻²)` = sprintf("%.2f ± %.2f",
                                          winter_adj_total_C_budget,
                                          se_winter_adj_total_C_budget)
  ) %>%
  select(year, winter_days, 
         `NEE (μmol m⁻² s⁻¹)`, `FCH4 (nmol m⁻² s⁻¹)`,
         `CO2 Budget (g C m⁻²)`, `CH4 Budget (g C m⁻²)`, 
         `Total C Budget (g C m⁻²)`)

print(combined_table)

# # Add to seasonal_budgets_winteryear
# 
# 
# # Select relevant columns to merge
# winter_adj_to_merge <- winter_adj_budgets %>%
#   select(year, 
#          # Daily fluxes
#          winter_adj_daily_NEE_flux, se_winter_adj_daily_NEE_flux,
#          winter_adj_daily_FCH4_flux, se_winter_adj_daily_FCH4_flux,
#          # Seasonal budgets
#          winter_adj_CO2_budget, se_winter_adj_CO2_budget,
#          winter_adj_CH4_budget, se_winter_adj_CH4_budget,
#          winter_adj_total_C_budget, se_winter_adj_total_C_budget)
# 
# # Merge with existing seasonal budgets
# seasonal_budgets_winteryear_adj <- seasonal_budgets_winteryear %>%
#   left_join(winter_adj_to_merge, by = "year") %>%
#   mutate(
#     # Set to NA for non-winter seasons
#     across(starts_with("winter_adj"), 
#            ~ifelse(season == "Winter", .x, NA))
#   )
# 



# Create formatted table


library(knitr)
library(kableExtra)



pub_table <- winter_adj_budgets %>%
  mutate(
    # Format with mean ± SE
    `Daily NEE Flux` = sprintf("%.3f ± %.3f", 
                               winter_adj_daily_NEE_flux, 
                               se_winter_adj_daily_NEE_flux),
    `Daily FCH4 Flux` = sprintf("%.4f ± %.4f",
                                winter_adj_daily_FCH4_flux,
                                se_winter_adj_daily_FCH4_flux),
    `CO2 Budget` = sprintf("%.2f ± %.2f",
                           winter_adj_CO2_budget,
                           se_winter_adj_CO2_budget),
    `CH4 Budget` = sprintf("%.3f ± %.3f",
                           winter_adj_CH4_budget,
                           se_winter_adj_CH4_budget),
    `Total C Budget` = sprintf("%.2f ± %.2f",
                               winter_adj_total_C_budget,
                               se_winter_adj_total_C_budget),
    `Winter Year` = year,
    `Days` = winter_days
  ) %>%
  select(`Winter Year`, Days, 
         `Daily NEE Flux`, `Daily FCH4 Flux`,
         `CO2 Budget`, `CH4 Budget`, `Total C Budget`)



# Statistical notes about SE propagation



#SE PROPAGATION METHODOLOGY\n")

#Standard errors were propagated as follows:\n\n")

#1. Daily Flux (μmol/m²/s or nmol/m²/s):\n")
#   Mean: flux_winter = flux_growing × proportion\n")
#   SE:   SE_winter = SE_growing × proportion\n\n")

#2. Daily Budget (g C/m²/day):\n")
#   Mean: budget_daily = flux_winter × conversion\n")
#   SE:   SE_daily = SE_flux × conversion\n\n")

#3. Seasonal Budget (g C/m²):\n")
#   Mean: budget_seasonal = budget_daily × n_days\n")
#   SE:   SE_seasonal = SE_daily × sqrt(n_days)\n")
#   (Assumes daily values are independent)\n\n")

#4. Total C Budget:\n")
#   Mean: total_C = CO2_budget + CH4_budget\n")
#   SE:   SE_total = sqrt(SE_CO2² + SE_CH4²)\n")
#   (Variance addition rule for independent variables)\n\n")



cat("Summary of winter-adjusted values:\n")
cat(sprintf("  Mean winter NEE flux: %.3f ± %.3f μmol m⁻² s⁻¹\n",
            mean(winter_adj_budgets$winter_adj_daily_NEE_flux),
            mean(winter_adj_budgets$se_winter_adj_daily_NEE_flux)))
cat(sprintf("  Mean winter FCH4 flux: %.4f ± %.4f nmol m⁻² s⁻¹\n",
            mean(winter_adj_budgets$winter_adj_daily_FCH4_flux),
            mean(winter_adj_budgets$se_winter_adj_daily_FCH4_flux)))
cat(sprintf("  Mean winter CO2 budget: %.2f ± %.2f g C m⁻²\n",
            mean(winter_adj_budgets$winter_adj_CO2_budget),
            mean(winter_adj_budgets$se_winter_adj_CO2_budget)))
cat(sprintf("  Mean winter CH4 budget: %.3f ± %.3f g C m⁻²\n\n",
            mean(winter_adj_budgets$winter_adj_CH4_budget),
            mean(winter_adj_budgets$se_winter_adj_CH4_budget)))
```
#save winter-adj tables

```{r}
#save winter-adj daily NEE and FCH4_F fluxes raw table 
 library(writexl)
write_xlsx(winter_adj_budgets, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/seasonal_means_budgets_winteryear_adj_12.17.2025.xlsx")


#save winter-adj daily NEE and FCH4_F fluxes formatted 

write_xlsx(combined_table, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/seasonal_means_budgets_winteryear_adj_formatted_12.17.2025.xlsx")

```

# Annual means and budgets 

##### (not winter-adj - did winter adj here manually in excel sheet "Seasonal_Cbudgets_Table_all_years_for_paper")

#Code for calculating budgets based on all seasons & assigned winter year 
```{r}
annual_budgets_winteryear <- df_avg2 %>%
  group_by(winter_year) %>%
  summarize(
   # pre-winter adj CO2 budget
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
   #pre-winter adj CH4 budget 
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
   #pre-winter adj total C budget
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    # GPP budget
    GPP_budget = sum(GPP_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
   # pre-winter adj RECO budget
    RECO_budget = sum(RECO_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    #measured soil temp, no GF
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
   #SWC_3, all 
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    #SWC_3, excluding winter and TS_3_gf > 2 degC
    mean_soil_moisture_nf = mean(SWC_3_nf, na.rm = TRUE),
    se_soil_moisture_nf = sd(SWC_3_nf, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_nf))),
   #measured TA, no GF
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
   #ERA-mod GF TA
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
    se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
   #ERA-mod GF TS_3
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled)))),
   #vapor pressure deficit
    mean_VPD = mean(VPD, na.rm = TRUE),
    se_VPD = sd(VPD, na.rm = TRUE) / sqrt(sum(!is.na(VPD))),
   #relative humidity 
    mean_RH = mean(RH, na.rm = TRUE),
    se_RH = sd(RH, na.rm = TRUE) / sqrt(sum(!is.na(RH))),
    # Sensible heat
    mean_H = mean(H, na.rm = TRUE),
    se_H = sd(H, na.rm = TRUE) / sqrt(sum(!is.na(H))),
    # Latent heat
    mean_LE = mean(LE, na.rm = TRUE),
    se_LE = sd(LE, na.rm = TRUE) / sqrt(sum(!is.na(LE))),
    # Ground heat
    mean_G = mean(G_1_1_1, na.rm = TRUE),
    se_G = sd(G_1_1_1, na.rm = TRUE) / sqrt(sum(!is.na(G_1_1_1))),

   #days per season 
   days = n(),
    .groups = "drop"
  ) %>%
  rename(year = winter_year) 

print(annual_budgets_winteryear)
```

#Make table with annual winter-year budgets - not winter-adj **this is by winter year, not calendar year, so the days don't match up to 365

```{r}
library(tidyverse)


#Create formatted columns with mean ± SE

annual_budgets_winteryear_formatted <- annual_budgets_winteryear %>%
 # arrange(year) %>%  # Sort by this order
  mutate(
    # Carbon budgets (no SE, just values)
    `CO2 Budget` = sprintf("%.2f", CO2_budget),
    `CH4 Budget` = sprintf("%.2f", CH4_budget),
    `Total C Budget` = sprintf("%.2f", total_C_budget),
     `GPP Budget` = sprintf("%.2f", GPP_budget),
     `RECO Budget` = sprintf("%.2f", RECO_budget),
    
    # Temperature variables
    `Soil Temp` = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    `Air Temp` = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    `Air Temp (GF)` = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    `Soil Temp (GF)` = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
    
    # Soil moisture
    `SWC` = sprintf("%.3f ± %.3f", mean_soil_moisture, se_soil_moisture),
    `SWC (non-frozen)` = sprintf("%.3f ± %.3f", mean_soil_moisture_nf, se_soil_moisture_nf),
    
    # Other met variables
    `VPD` = sprintf("%.2f ± %.2f", mean_VPD, se_VPD),
    `RH` = sprintf("%.2f ± %.2f", mean_RH, se_RH),
    `H` = sprintf("%.2f ± %.2f", mean_H, se_H),
    `LE` = sprintf("%.2f ± %.2f", mean_LE, se_LE),
    `G` = sprintf("%.2f ± %.2f", mean_G, se_G),
    
    # Keep identifiers and sample size
    Year = year,
    Days = days
  ) %>%
  # Select formatted columns in desired order
  select(Year, Days, 
         `CO2 Budget`, `CH4 Budget`, `Total C Budget`, `GPP Budget`, `RECO Budget`,
         `Air Temp (GF)`, `Soil Temp (GF)`,
         `SWC (non-frozen)`,
         `VPD`, `RH`, `H`, `LE`, `G`)

# Display as simple table
print(annual_budgets_winteryear_formatted)



```


#Save table 
```{r}
# # Save to Excel
 library(writexl)
write_xlsx(annual_budgets_winteryear_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/annual_means_budgets_winteryear_12.17.2025.xlsx")
```

# % gapfilled - edit 
#### comparing HH and df_avg %gapfilled differ by only ~1%, so using df_avg for calculating gf%

```{r}

gapfill_summary <- df_avg2 %>%
  # Just group directly by winter_year (which already exists!)
  group_by(winter_year) %>%
  summarise(
    # Total days in this season
    total_days = n(),
    
    # Air Temperature (TA vs TA_gapfilled)
    TA_measured = sum(!is.na(TA)),
    TA_gapfilled = sum(is.na(TA) & !is.na(TA_gapfilled)),
    TA_pct_gapfilled = round((TA_gapfilled / total_days) * 100, 1),
    
    # Soil Temperature (TS_3_1_1 vs TS_3_gapfilled)
    TS_measured = sum(!is.na(TS_3_1_1)),
    TS_gapfilled = sum(is.na(TS_3_1_1) & !is.na(TS_3_gapfilled)),
    TS_pct_gapfilled = round((TS_gapfilled / total_days) * 100, 1),
    
    .groups = "drop"
  ) %>%
  # Rename to "year" for consistency with other tables
  rename(year = winter_year)



# Display summary
print(gapfill_summary)


#compact table with only %gf

gapfill_pct_only <- gapfill_summary %>%
  select(year, total_days,
         TA_pct_gapfilled,
         TS_pct_gapfilled)

print(gapfill_pct_only)


```

#merge %gf with seasonal summaries by winter_year
```{r}

seasonal_budgets_winteryear_with_gf <- seasonal_budgets_winteryear %>%
  left_join(
    gapfill_pct_only,
    by = c("year", "season")
  )


# Display merged data
print(seasonal_budgets_winteryear_with_gf %>%
      select(year, season, days, 
             mean_daily_NEE, NEE_pct_gapfilled,
             mean_daily_FCH4, FCH4_pct_gapfilled,
             ))

#formatted table 

seasonal_table_with_gf <- seasonal_budgets_winteryear_with_gf %>%
  mutate(
    # Format mean ± SE with % gapfilled
    `NEE` = sprintf("%.2f ± %.2f (%s%%)", 
                    mean_daily_NEE, se_NEE, NEE_pct_gapfilled),
    `FCH4` = sprintf("%.2f ± %.2f (%s%%)", 
                     mean_daily_FCH4, se_FCH4, FCH4_pct_gapfilled),
    `GPP` = sprintf("%.2f ± %.2f (%s%%)",
                    mean_daily_GPP, se_GPP, GPP_pct_gapfilled),
    `RECO` = sprintf("%.2f ± %.2f (%s%%)",
                     mean_daily_RECO, se_RECO, RECO_pct_gapfilled),
    `Air Temp` = sprintf("%.2f ± %.2f (%s%%)",
                         mean_TA_gapfilled, se_TA_gapfilled, TA_pct_gapfilled),
    `Soil Temp` = sprintf("%.2f ± %.2f (%s%%)",
                          mean_TS_3_gapfilled, se_TS_3_gapfilled, TS_pct_gapfilled),
    
    Year = year,
    Season = season,
    Days = days
  ) %>%
  select(Year, Season, Days, NEE, FCH4, GPP, RECO, `Air Temp`, `Soil Temp`)

seasonal_table_with_gf 

```
#save table with %gf
```{r}

 library(writexl)
write_xlsx(seasonal_table_with_gf , "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/seasonal_means_budgets_winteryear_percentGF_12.17.2025.xlsx")

```



#end 
```{r}

```
















```{r}
# CH4 in nano-moles of CH4/m^2/s)
#Net CH4 Flux - gapfilled

# Calculate 0.443 (~44%) of daily average CH4 flux (FCH4_F) during growing season season with standard error

seasonal_FCH4_summary_gf_winteradjCH4 <- df_avg2 %>%
  filter(season == "Growing Season") %>%
  group_by(year, season) %>%
  summarize(
    mean_flux = mean(FCH4_F*0.443, na.rm = TRUE),
    se_flux = sd(FCH4_F*0.443, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
    n_days = n()  # number of days per season
  )

#renaming the ~44% of growing season avg variable name 
seasonal_FCH4_summary_gf_winteradjCH4$season[seasonal_FCH4_summary_gf_winteradjCH4$season=="Growing Season"] <- "Winter Adj FCHF DailyAvg"


seasonal_FCH4_summary_gf_winteradjCH4 

# Format the results with mean ± SE
seasonal_FCH4_summary_gf_winteradjCH4_formatted <- seasonal_FCH4_summary_gf_winteradjCH4 %>%
  mutate(
    flux = sprintf("%.2f ± %.2f", mean_flux, se_flux)
  ) %>%
  select(season, flux)
         
seasonal_FCH4_summary_gf_winteradjCH4_formatted

# # Save to Excel
library(writexl)
#write_xlsx(seasonal_FCH4_summary_gf_winteradjCH4_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteradj_dailyAvg_FCH4_summary_gf.xlsx")

```


# NEE - daily flux winter-adj

#based on 2018 dailg avg growing season fluxes, winter daily avg CO2 flux can be ~45% of growing season daily avg flux (0.451) --> multiply winter daily avg fluxes by 0.451 & re-calc the daily avg flux to get an adjusted potential winter daily avg flux 


```{r}
#FC in umol/m2/s

# Calculate 0.473 (~47%) of daily average daily NEE (FC_F) during growing season season with standard error

seasonal_NEE_summary_gf_winteradjNEE <- df_avg %>%
  filter(season == "Growing Season") %>%
  group_by(year, season) %>%
  summarize(
    mean_flux = abs(mean(FC_F*0.451, na.rm = TRUE)),
    se_flux = sd(FC_F*0.451, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
    n_days = n()  # number of days per season
  )

#renaming the ~47% of growing season avg variable name 
seasonal_NEE_summary_gf_winteradjNEE$season[seasonal_NEE_summary_gf_winteradjNEE$season=="Growing Season"] <- "Winter Adj NEE DailyAvg"


seasonal_NEE_summary_gf_winteradjNEE 

# Format the results with mean ± SE
seasonal_NEE_summary_gf_winteradjNEE_formatted <- seasonal_NEE_summary_gf_winteradjNEE  %>%
  mutate(
    flux = sprintf("%.2f ± %.2f", mean_flux, se_flux)
  ) %>%
  select(season, flux)
         
seasonal_NEE_summary_gf_winteradjNEE_formatted

# # Save to Excel
# library(writexl)
#write_xlsx(seasonal_NEE_summary_gf_winteradjNEE_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteradj_dailyAvg_NEE_summary_gf.xlsx")


```







