---
title: "Council CH4 and temp / moisture" #playing around with empirical model using soil temp and soil moisture for methane 
output: html_document
date: "2024-12-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)


Sys.setenv(TZ='UTC')
```

# Set working directory and then Load data

```{r}
#original half-hourly dataframe 
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


#needs continuous day variable
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

#needs continuous day variable
df_monthly_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_monthly_AVG_gapfilled_clean_2017_2023_for analysis.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


#Notes:
#SWC
#SWC also broken up by location
#  SWC_1_1_1 % Soil water content (15cm depth) – margin pond
# SWC_2_1_1 % Soil water content (15cm depth) – lichen/berries
# SWC_3_1_1 % Soil water content (15cm depth) - tussock ***** switch to using this one ** (used SWC1 before, by margin pond, switching to SWC3, tussock)

#Temp
#quick look at the different temp profiles (all at 15 cm depth, diff locations)
# sum(is.na(df$TS_1_1_1)) #60277 NA's in temp by margin pond 
# sum(is.na(df$TS_2_1_1)) #48729 NA's in temp  by lichen/berries  
# sum(is.na(df$TS_3_1_1)) #50445 NA's in temp by tussock 
# sum(is.na(df$TS_4_1_1)) #52796 NA's in temp by foot of EC tower
# sum(is.na(df$TS_5_1_1)) #new temp in latest upload with 2023 data, don't have info on what it is yet** 


```


#calculate the correlations between FCH4, TS_3_1_1, and SWC_3_1_1:
#### correlation coefficients explain the strength and direction of a correlation, but not the magnitude of change (like the slope from an lm analysis does) and not the goodness of fit of the model (R2)
```{r}
#these TS and SWC are measurements from tussock area of site 

# Calculate correlations of FCH4 and soil temp, SWC
cor(df$FCH4, df$TS_3_1_1, use = "complete.obs") #uses only rows with no missing values in any of the variables being correlated, as specified by the equation 
cor(df$FCH4, df$SWC_3_1_1, use = "complete.obs")

# Interaction term of soil temp and SWC 
df$interaction_tuss <- df$TS_3_1_1 * df$SWC_3_1_1
cor(df$FCH4, df$interaction_tuss, use = "complete.obs")

#Results: no strong correlations - all very weak pos but interaction a bit better of a fit 

# [1] 0.1981594 FCH4 vs temp
# [1] 0.153488 FCH4 vs SWC
# [1] 0.2642889 FCH4 vs interaction of temp and SWC


#Now testing corr between temp and SWC from other locations: 


#measurements from margin pond for both temp and SWC 

# Calculate correlations of FCH4 and soil temp, SWC
cor(df$FCH4, df$TS_1_1_1, use = "complete.obs") 
cor(df$FCH4, df$SWC_1_1_1, use = "complete.obs")

# Interaction term of soil temp and SWC 
df$interaction_pond <- df$TS_1_1_1 * df$SWC_1_1_1
cor(df$FCH4, df$interaction_pond, use = "complete.obs")
#Results:
# [1] 0.2358536
# [1] 0.232435
# [1] 0.3072315 --> corr best with interaction btwn temp and SWC


#measurements from lichen/berries for both temp and SWC 

# Calculate correlations of FCH4 and soil temp, SWC
cor(df$FCH4, df$TS_2_1_1, use = "complete.obs") 
cor(df$FCH4, df$SWC_2_1_1, use = "complete.obs")

# Interaction term of soil temp and SWC 
df$interaction_liBer <- df$TS_2_1_1 * df$SWC_2_1_1
cor(df$FCH4, df$interaction_liBer, use = "complete.obs")
#Results:
# [1] 0.1786063
# [1] 0.1145262
# [1] 0.2269394 #best with interaction effect 

#Note: best correlation between FCH4 and temp/SWC came from margin pond location, and included the interaction of temp and SWC, but corr = 0.307

```

#Plot to visualize 
```{r}
# Scatter plots of correlations - tussock 
ggplot(df, aes(x = TS_3_1_1, y = FCH4)) + geom_point() + geom_smooth(method = "lm")
ggplot(df, aes(x = SWC_3_1_1, y = FCH4)) + geom_point() + geom_smooth(method = "lm")
ggplot(df, aes(x = interaction, y = FCH4)) + geom_point() + geom_smooth(method = "lm") 


# Scatter plots of correlations - pond margin 
ggplot(df, aes(x = TS_1_1_1, y = FCH4)) + geom_point() + geom_smooth(method = "lm")
ggplot(df, aes(x = SWC_1_1_1, y = FCH4)) + geom_point() + geom_smooth(method = "lm")
ggplot(df, aes(x = interaction_pond, y = FCH4)) + geom_point() + geom_smooth(method = "lm") 

# Scatter plots of correlations - lichen and berries 
ggplot(df, aes(x = TS_2_1_1, y = FCH4)) + geom_point() + geom_smooth(method = "lm")
ggplot(df, aes(x = SWC_2_1_1, y = FCH4)) + geom_point() + geom_smooth(method = "lm")
ggplot(df, aes(x = interaction_liBer, y = FCH4)) + geom_point() + geom_smooth(method = "lm") 

```
#lm of relationshipsfrom dataset; testing all soil and SWC locations 
```{r}
#df without NAs 
df_clean <- na.omit(df[, c("FCH4", "TS_3_1_1", "TS_2_1_1", "TS_1_1_1" )])

#these results of using df with or without NA's = same stats 
model1 <-lm(FCH4 ~ TS_3_1_1, data = df_clean)
summary(model1) 

#temp - tussock
model2 <-lm(FCH4 ~ TS_3_1_1, data = df)
summary(model1) #slope = 0.48; R2 = 0.06, p <0.001; Residual standard error: 9.259

#temp - lichen and berries 
model3 <-lm(FCH4 ~ TS_2_1_1, data = df)
summary(model3) #slope = 0.36; R2 = 0.03, p <0.001; Residual standard error: 9.453

#temp - pond margin
model4 <-lm(FCH4 ~ TS_1_1_1, data = df)
summary(model4) #slope = 0.35; R2 = 0.05; p <0.001; Residual standard error: 9.304 

#SWC - tussock
model5 <-lm(FCH4 ~ SWC_3_1_1, data = df)
summary(model5) #slope = 0.07; R = 0.02; p <0.001; Residual standard error: 9.474 

#SWC - lichen and berries
model6 <-(lm(FCH4 ~ SWC_2_1_1, data = df))
summary(model6) #slope = 0.06; R2 = 0.01; p<0.001; Residual standard error: 9.543

#SWC - pond margin 
model7 <-lm(FCH4 ~ SWC_1_1_1, data = df)
summary(model7) #slope = 0.12; R2 = 0.05; p<0.001; Residual standard error: 9.312

#interaction - tussock 
model8 <-lm(FCH4 ~ interaction_tuss, data = df)
summary(model8) #slope = 0.0088; R2 = 0.0698; p<0.001; Residual standard error: 9.247 

#interaction - pond margin 
model9 <-lm(FCH4 ~ interaction_pond, data = df)
summary(model9) #slope = 0.009; R2 = 0.09; p<0.001; Residual standard error: 9.111

#interaction - lichen and berries 
model10 <-lm(FCH4 ~ interaction_liBer, data = df)
summary(model10) #slope = 0.01; R2 = 0.05; p<0.001; Residual standard error: 9.357



model11 <-lm(FCH4 ~ TS_3_1_1 + SWC_3_1_1 + interaction_tuss, data = df)
summary(model11) #R2 = 0.11; p<0.001

model12 <-lm(FCH4 ~ TS_2_1_1 + SWC_2_1_1 + interaction_liBer, data = df)
summary(model12) #R2 = 0.06, p<0.001

model13 <-lm(FCH4 ~ TS_1_1_1 + SWC_1_1_1 + interaction_pond, data = df)
summary(model13) #R2 = 0.11, p<0.001


#TO DO, look at RSMEs again

```


#predictive model, trying linear model as starting point 
```{r}
# Linear model - soil temp, SWC, and interaction 
model <- lm(FCH4 ~ TS_3_1_1 + SWC_3_1_1 + interaction, data = df)
summary(model)

# Predict missing values
df$predicted_FCH4_lm <- predict(model, newdata = df)

#Results:
# TS_3_1_1:  -1.4649, as temp inc, methane dec? p= (<2e-16) - SIG  **not expecting that result, esp since plot shows pos trend..? inc temp should inc methane emissions # SWC_3_1_1: The coefficient for soil moisture is -0.0001 and p=0.9,NOT sig 
# interaction: 0.0352, R2 = 0.011 so pos correlation but very weak; but p  <2e-16 - SIG*

# Linear model 2 - soil temp only 
model2 <- lm(FCH4 ~ TS_3_1_1, data = df)
summary(model2)

#Results: 
#TS_3_1_1     est =0.35596    sE=0.01120   p= <2e-16 ***; Multiple R-squared:  0.03927,	Adjusted R-squared:  0.03923 
# Predict missing values
df$predicted_FCH4.2 <- predict(model2, newdata = df)

# Linear model 3 - interaction only 
model3 <- lm(FCH4 ~ interaction, data = df)
summary(model3)

#Results: 
#interaction est=0.0088479  SE=0.0002053   p= <2e-16 ***; Multiple R-squared:  0.06985,	Adjusted R-squared:  0.06981 

# Predict missing values
df$predicted_FCH4.3 <- predict(model3, newdata = df)


# Linear model - soil temp and interaction 
model4 <- lm(FCH4 ~ TS_3_1_1 + interaction, data = df)
summary(model)

# Predict missing values
df$predicted_FCH4.4 <- predict(model4, newdata = df)
#Results: same as model(1)
#Multiple R-squared:  0.1161,	Adjusted R-squared:  0.116 
#F-statistic:  1083 on 3 and 24733 DF,  p-value: < 2.2e-16

```


#RMSE of various linear model options
####evaluate the model using RMSE(measures the magnitude of errors between predicted and actual values; lower is better fit of model to data)
```{r}
# Calculate RMSE - lower RMSE = better fit 

#soil temp, SWC, and interaction --> slightly better than the other two
rmse <- sqrt(mean((df$FCH4 - df$predicted_FCH4_lm)^2, na.rm = TRUE))
print(paste("RMSE: ", rmse)) # "RMSE:  9.01373878828402"

#soil temp only 
rmse <- sqrt(mean((df$FCH4 - df$predicted_FCH4.2)^2, na.rm = TRUE))
print(paste("RMSE: ", rmse)) #"RMSE:  9.39753297747576"

#interaction only 
rmse <- sqrt(mean((df$FCH4 - df$predicted_FCH4.3)^2, na.rm = TRUE))
print(paste("RMSE: ", rmse)) #"RMSE:  9.24675505010849"

#soil temp and interaction
rmse <- sqrt(mean((df$FCH4 - df$predicted_FCH4.4)^2, na.rm = TRUE))
print(paste("RMSE: ", rmse)) #"RMSE:  9.01373897029865" --> same as first lm model --> confirms SWC doesn't sig impact model, therefore not running model with just SWC/don't need SWC in this model 

```
#If linear models don't work well, try random forest but use interaction effect? 
```{r}
# This chunk is useful for getting an initial sense of how well the Random Forest model performs on the entire dataset. However, it doesn't provide a robust evaluation since it uses the same data for both training and prediction, which can lead to overfitting and therefore also a RMSE that looks better than it should be - that is why this output seems to perform so well, but shouldn't be trusted until trained on a portion of the df and tested on the other portion 

# Load necessary library
library(randomForest)

# set up random forest model 
rf_model <- randomForest(FCH4 ~ TS_3_1_1 + SWC_3_1_1 + interaction, data = df, na.action = na.omit)
df$predicted_FCH4_rf <- predict(rf_model, newdata = df)

# Calculate RMSE
rmse_rf <- sqrt(mean((df$FCH4 - df$predicted_FCH4_rf)^2, na.rm = TRUE))
print(paste("RMSE (Random Forest): ", rmse_rf)) # "RMSE (Random Forest):  4.54702260438728"  --> improved RMSE by a lot but this is not on tested data, so ignore this chunk and its results 
```
#Train new random forest model 
```{r}
# This chunk splits the dataset into training and testing sets, trains the Random Forest model on the training set, and then evaluates its performance on the test set. It calculates both RMSE and R-squared to assess the model's accuracy. *Produces a higher but more accurate RMSE and R-squared value 

#*note, output here is "predicted_CH4_RF" 

# Load necessary libraries
library(tidyverse)
library(randomForest)
library(caret)


# Create the interaction term
df$interaction <- df$TS_3_1_1 * df$SWC_3_1_1

# Drop rows with NA in FCH4
df <- df %>% drop_na(FCH4)
df <- df %>% drop_na(SWC_3_1_1)
df <- df %>% drop_na(TS_3_1_1)
df <- df %>% drop_na(interaction)

# Split the data into training and testing sets
set.seed(123) #seed 123 is what we use in our processing code 
trainIndex <- createDataPartition(df$FCH4, p = .8, list = FALSE)
trainData <- df[trainIndex,]
testData <- df[-trainIndex,]

# Train a Random Forest model
rf_model_training <- randomForest(FCH4 ~ TS_3_1_1 + SWC_3_1_1 + interaction, data = trainData, na.action = na.omit)

# Predict on the test set
testData$predicted_FCH4_RF <- predict(rf_model_training, newdata = testData)

# Calculate RMSE between the true data and predicted data
rmse_rf <- sqrt(mean((testData$FCH4 - testData$predicted_FCH4_RF)^2))
print(paste("RMSE (Random Forest): ", rmse_rf))
#Results:  "RMSE (Random Forest):  7.96643546914752"--> does improve model a bit compared to just the lm 

# Calculate R-squared
r2_rf <- cor(testData$FCH4, testData$predicted_FCH4_RF)^2
print(paste("R-squared (Random Forest): ", r2_rf))
#Results: "R-squared (Random Forest):  0.304920313430615"
```

#Comparing RF modeled data to actual data 
```{r}
# Compare actual vs predicted values
comparison <- testData %>% select(FCH4, predicted_FCH4_RF)
print(head(comparison))
#don't compare very well....



summary(lm(testData$FCH4 ~ testData$predicted_FCH4_RF))
#Results: not a great fit 
#slope = 0.85; Multiple R-squared:  0.3049,	Adjusted R-squared:  0.3048 
#F-statistic:  2168 on 1 and 4943 DF,  p-value: < 2.2e-16

# Plot the comparison
ggplot(comparison, aes(x = FCH4, y = predicted_FCH4_RF)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted Methane Flux (FCH4)",
       x = "Actual FCH4",
       y = "Predicted FCH4 from RF training") +
   annotate(geom = 'text',x = 50, y = -10,label=expression(R^2~"= 0.3"),size = 3)+ 
     annotate(geom = 'text',x = 50, y = -15,label= "slope = 0.85",size = 3)+ 
  theme_minimal()


```
#comparing predicted CH4 from RF and lm from test Data to actual CH4 data 
```{r}
#FCH4 vs FCH4 from RF, looking at comparison of test data 
summary(lm(testData$FCH4 ~ testData$predicted_FCH4_RF))
#Results: not a great fit 
#slope = 0.85; Multiple R-squared:  0.3049,	Adjusted R-squared:  0.3048 
#F-statistic:  2168 on 1 and 4943 DF,  p-value: < 2.2e-16

# Plot the comparison of CH4 and predicted_FCH4_rf from random forest 
ggplot(testData, aes(x = FCH4, y = predicted_FCH4_RF)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted Methane Flux (FCH4)",
       x = "Actual FCH4",
       y = "Predicted FCH4_rf") +
    annotate(geom = 'text',x = 50, y = -10,label=expression(R^2~"= 0.3"),size = 3)+ 
     annotate(geom = 'text',x = 50, y = -15,label= "slope = 0.85",size = 3)+ 
  annotate(geom = 'text',x = 57, y = -10,label= "RMSE = 7.97",size = 3)+ 
  theme_minimal()


#testing actual vs predicted CH4 from original df 
model5 <- lm(FCH4 ~ predicted_FCH4_lm, data = df)
summary(model5)
# Results: slope = 1; Multiple R-squared:  0.1161,	Adjusted R-squared:  0.1161 
#F-statistic:  3250 on 1 and 24735 DF,  p-value: < 2.2e-16


# Plot the comparison of actual CH4 and predicted CH4 from lm from orig df
ggplot(df, aes(x = FCH4, y = predicted_FCH4_lm)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted Methane Flux (FCH4)",
       x = "Actual FCH4",
       y = "Predicted FCH4_LM") +
     annotate(geom = 'text',x = 50, y = 0,label=expression(R^2~"= 0.12"),size = 3)+ 
     annotate(geom = 'text',x = 50, y = 1,label= "slope = 1",size = 3)+ 
   annotate(geom = 'text',x = 50, y = -1,label= "RMSE = 9.01",size = 3)+ 
  theme_minimal()


#Need to merge df before this can work 
# Plot the comparison of predicted CH4_lm and predicted_FCH4_rf
# ggplot(df, aes(x = predicted_FCH4_lm, y = predicted_FCH4_RF)) +
#   geom_point(alpha = 0.3) +
#   geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
#   labs(title = "Actual vs Predicted Methane Flux (FCH4)",
#        x = "predicted FCH4",
#        y = "Predicted FCH4_rf") +
#   theme_minimal()

#CH4 data vs predicted FCH4_lm from the testdata 
ggplot(data = testData,aes(FCH4, predicted_FCH4_lm))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1) +
  annotate(geom = 'text',x = 50, y = 0,label=expression(R^2~"= 0.12"),size = 3)+ 
     annotate(geom = 'text',x = 50, y = 1,label= "slope = 1",size = 3)+ 
   annotate(geom = 'text',x = 50, y = -1,label= "RMSE = 9.01",size = 3) # +
    # scale_x_continuous(limits = c(-50,150),expression('Random Forest predicted FCH4  ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
    # scale_y_continuous(limits = c(-50,150),expression('Eddy Covariance FCH4 ('*mu*mol~CO[2]~m^-2~s^-1*')'))#+
  # annotate(geom = 'text',x = 50, y = -8,label=expression(R^2~"= 0.47"),size = 3)+
  # annotate(geom = 'text',x = 50,y = -15,label=expression(Slope~"= 1.01"),size = 3)+
  # theme(text = element_text(size = 8))

  #FCH4 data vs predicted FCH4_RF from testData 
  ggplot(data = testData,aes(FCH4, predicted_FCH4_RF))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1) +
    annotate(geom = 'text',x = 50, y = -10,label=expression(R^2~"= 0.3"),size = 3)+ 
     annotate(geom = 'text',x = 50, y = -15,label= "slope = 0.85",size = 3)+ 
  annotate(geom = 'text',x = 65, y = -10,label= "RMSE = 7.97",size = 3)

```
















