---
title: "Alt Methane Model budgets" #examining C budgets with these diferent methane models 
output: html_document
date: "2025-01-22"
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)


Sys.setenv(TZ='UTC')
```

# Set working directory and then Load data

TO DO: re-running all these with SWC_3 (tussock) version of dataset - KK 1/7/2025 **use the era dataset** 
```{r}
df = fread(input ="C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_HH_alt_methane_models.csv", na.strings = "-9999")


colnames(df)

summary(df$date)

```

#Create daily avg for ease
```{r}
library(openair)
#average entire dataframe with tighter data availability threshold to avoid using large data gaps in calculating the avg -- having at least 50% of the data 
df_avg <- as.data.frame(date)
df_avg <-timeAverage(df, avg.time = "day", data.thresh = 50) #this seems to remove the measured fch4, probably due to severe lack of data 



# Create seasonal assignments using both year and DOY
df_avg <- df_avg %>%
  mutate(
    season = case_when(
      # 2017 seasons
      year == "2017" & (DOY >= 258 & DOY <= 290) ~ 'Fall Senescence',
      year == "2017" & (DOY >= 131 & DOY <= 257) ~ 'Growing Season',
      year == "2017" & (DOY >= 291 | DOY <= 130) ~ 'Winter',
      
      
      # 2018 seasons
      year == "2018" &  (DOY >= 257 & DOY <= 300) ~ 'Fall Senescence',
      year == "2018" & (DOY >= 152 & DOY <= 256) ~ 'Growing Season',
      year == "2018" & (DOY >= 301 | DOY <= 151) ~ 'Winter',
      
      # 2019 seasons
      year == "2019" & (DOY >= 246 & DOY <= 284) ~ 'Fall Senescence',
      year == "2019" & (DOY >= 145 & DOY <= 245) ~ 'Growing Season',
      year == "2019" & (DOY >= 285 | DOY <= 144) ~ 'Winter',
      
      # 2020 seasons
      year == "2020" & (DOY >= 252 & DOY <= 299) ~ 'Fall Senescence',
      year == "2020" & (DOY >= 136 & DOY <= 251) ~ 'Growing Season',
      year == "2020" & (DOY >= 300 | DOY <= 135) ~ 'Winter',
      
      # 2021 seasons
      year == "2021" & (DOY >= 260 & DOY <= 283) ~ 'Fall Senescence',
      year == "2021" &(DOY >= 142 & DOY <= 259) ~ 'Growing Season',
      year == "2021" &(DOY >= 284 | DOY <= 141) ~ 'Winter',
      
      # 2022 seasons
      year == "2022" & (DOY >= 245 & DOY <= 281) ~ 'Fall Senescence',
      year == "2022" & (DOY >= 144 & DOY <= 244) ~ 'Growing Season',
      year == "2022" & (DOY >= 282 | DOY <= 143) ~ 'Winter',
      
      # not doing one for 2023 ** incomplete year 
      TRUE ~ NA_character_
    )
  )


```



####Yearly df 
```{r}
df_avg_2017 <- df_avg %>%
  filter(year(date) == 2017)

df_avg_2018 <- df_avg %>%
  filter(year(date) == 2018)

df_avg_2019 <- df_avg %>%
  filter(year(date) == 2019)

df_avg_2020 <- df_avg %>%
  filter(year(date) == 2020)

df_avg_2021 <- df_avg %>%
  filter(year(date) == 2021)

df_avg_2022 <- df_avg %>%
  filter(year(date) == 2022)

df_avg_2023 <- df_avg %>%
  filter(year(date) == 2023)


#sum - yearly budget 
sum(df_avg_2017$rfch4_gf * 60 * 60 * 24 * (1/1000000000) * 12 )
sum(df_avg_2018$rfch4_gf * 60 * 60 * 24 * (1/1000000000) * 12 )

```
```{r}
convert_ch4_flux <- function(flux_nmol) {
  # Starting with nano-molCH4-C/m2/s
  # * 60*60*24 for seconds in a day
  # * (1/1000000000) for nano conversion
  # * 12 for atomic mass of C
  flux_nmol * 60 * 60 * 24 * (1/1000000000) * 12
}

# Calculate annual budgets
annual_budget <- df_avg %>%
  filter(year != 2023) %>%
  group_by(year) %>%
  summarise(
    XGBoost_total = sum(fch4_gapfilled) * 60 * 60 * 24 * (1/1000000000) * 12,
    XGBoost_interact_total = sum(fch4_gapfilled_interact) * 60 * 60 * 24 * (1/1000000000) * 12,
    RandomForest_total = sum(rfch4_gf) * 60 * 60 * 24 * (1/1000000000) * 12,
    n_days = n()
  )

# Calculate seasonal budgets
seasonal_budget <- df_avg %>%
  filter(year != 2023) %>%
  group_by(year, season) %>%
  summarise(
    XGBoost_total = sum(fch4_gapfilled) * 60 * 60 * 24 * (1/1000000000) * 12,
    XGBoost_interact_total = sum(fch4_gapfilled_interact) * 60 * 60 * 24 * (1/1000000000) * 12,
    RandomForest_total = sum(rfch4_gf) * 60 * 60 * 24 * (1/1000000000) * 12,
    n_days = n()
  )
```



```{r}
# # Function to convert from nano-mol/m2/s to g CH4-C/m2/day --> *60*60*24*(1/1000000000)*12*
# convert_ch4_flux <- function(flux_nmol) {
#   # flux_nmol is in nano-mol/m²/s
#   # Convert to g/m²/day:
#   # * (24*60*60) for seconds in day
#   # * 16.04 for molar mass of CH4 --> but just checking C, so substitute 12 
#   # * 10^-9 for nano to base units
#   flux_nmol * (24*60*60) * 12 * 10^-9
# }

# Create visualization of annual budgets
annual_plot <- ggplot(annual_budget %>%
                     pivot_longer(cols = contains("total"),
                                names_to = "Model",
                                values_to = "CH4_flux")) +
  geom_bar(aes(x = as.factor(year), y = CH4_flux, fill = Model),
           position = "dodge", stat = "identity") +
  labs(title = "Annual CH4 Budgets by Model",
       x = "Year",
       y = expression(paste("CH"[4], " (g ", m^-2, " ", yr^-1, ")"))) +
  theme_bw()

# Create visualization of seasonal budgets
seasonal_plot <- ggplot(seasonal_budget %>%
                       pivot_longer(cols = contains("total"),
                                  names_to = "Model",
                                  values_to = "CH4_flux")) +
  geom_bar(aes(x = as.factor(year), y = CH4_flux, fill = Model),
           position = "dodge", stat = "identity") +
  facet_wrap(~season) +
  labs(title = "Seasonal CH4 Budgets by Model",
       x = "Year",
       y = expression(paste("CH"[4], " (g ", m^-2, " ", season^-1, ")"))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display results and plots
print("Annual CH4 Budgets (g CH4/m²/year):")
print(annual_budget)

print("\nSeasonal CH4 Budgets (g CH4/m²/season):")
print(seasonal_budget)

annual_plot
seasonal_plot
```


#Compare original RF to current model outputs



```{r}
df_avg_orig = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))
```

#Seasonal delineations for orig RF model dataset

```{r}
# First create DOY from date
df_avg_orig$DOY <- as.numeric(format(df_avg_orig$date, "%j"))

# add year column if not already present
df_avg_orig <- df_avg_orig %>%
  mutate(year = format(date, "%Y"))


# Create seasonal assignments using both year and DOY
df_avg_orig <- df_avg_orig  %>%
  mutate(
    season = case_when(
      # 2017 seasons
      year == "2017" & (DOY >= 258 & DOY <= 290) ~ 'Fall Senescence',
      year == "2017" & (DOY >= 131 & DOY <= 257) ~ 'Growing Season',
      year == "2017" & (DOY >= 291 | DOY <= 130) ~ 'Winter',
      
      
      # 2018 seasons
      year == "2018" &  (DOY >= 257 & DOY <= 300) ~ 'Fall Senescence',
      year == "2018" & (DOY >= 152 & DOY <= 256) ~ 'Growing Season',
      year == "2018" & (DOY >= 301 | DOY <= 151) ~ 'Winter',
      
      # 2019 seasons
      year == "2019" & (DOY >= 246 & DOY <= 284) ~ 'Fall Senescence',
      year == "2019" & (DOY >= 145 & DOY <= 245) ~ 'Growing Season',
      year == "2019" & (DOY >= 285 | DOY <= 144) ~ 'Winter',
      
      # 2020 seasons
      year == "2020" & (DOY >= 252 & DOY <= 299) ~ 'Fall Senescence',
      year == "2020" & (DOY >= 136 & DOY <= 251) ~ 'Growing Season',
      year == "2020" & (DOY >= 300 | DOY <= 135) ~ 'Winter',
      
      # 2021 seasons
      year == "2021" & (DOY >= 260 & DOY <= 283) ~ 'Fall Senescence',
      year == "2021" &(DOY >= 142 & DOY <= 259) ~ 'Growing Season',
      year == "2021" &(DOY >= 284 | DOY <= 141) ~ 'Winter',
      
      # 2022 seasons
      year == "2022" & (DOY >= 245 & DOY <= 281) ~ 'Fall Senescence',
      year == "2022" & (DOY >= 144 & DOY <= 244) ~ 'Growing Season',
      year == "2022" & (DOY >= 282 | DOY <= 143) ~ 'Winter',
      
      # not doing one for 2023 ** incomplete year 
      TRUE ~ NA_character_
    )
  )


```



####Yearly df 
```{r}
df_avg_orig_2017 <- df_avg_orig %>%
  filter(year(date) == 2017)

df_avg_orig_2018 <- df_avg_orig %>%
  filter(year(date) == 2018)

df_avg_orig_2019 <- df_avg_orig %>%
  filter(year(date) == 2019)

df_avg_orig_2020 <- df_avg_orig %>%
  filter(year(date) == 2020)

df_avg_orig_2021 <- df_avg_orig %>%
  filter(year(date) == 2021)

df_avg_orig_2022 <- df_avg_orig %>%
  filter(year(date) == 2022)

df_avg_orig_2023 <- df_avg_orig %>%
  filter(year(date) == 2023)


#sum - yearly budget 
sum(df_avg_orig_2017$FCH4_F * 60 * 60 * 24 * (1/1000000000) * 12 )
sum(df_avg_orig_2018$FCH4_F * 60 * 60 * 24 * (1/1000000000) * 12 )

```


#Original RF model annual and seasonal budgets 
```{r}

# Calculate annual budgets for gapfilled methane only - in g/m2
annual_budget_orig <- df_avg_orig %>%
  filter(year != 2023) %>%
  group_by(year) %>%
  summarise(
    FCH4_F = sum(FCH4_F) * 60 * 60 * 24 * (1/1000000000) * 12,
    n_days = n()
  )

# Calculate seasonal budgets for gapfilled methane only - in g/m2
seasonal_budget_orig <- df_avg_orig %>%
  filter(year != 2023) %>%
  group_by(year, season) %>%
  summarise(
    FCH4_F = sum(FCH4_F) * 60 * 60 * 24 * (1/1000000000) * 12,
     n_days = n()
    )

#Print
print(annual_budget_orig)

print(seasonal_budget_orig)
```



```{r}
# Create visualization of annual budgets
annual_plot_orig <- ggplot(annual_budget_orig, aes(x = as.factor(year), y = FCH4_F)) +
  geom_bar(stat = "identity", fill = "salmon") +
  labs(title = "Annual CH4 Budgets - Original RF",
       x = "Year",
       y = expression(paste("CH"[4], " (g ", m^-2, " ", yr^-1, ")"))) +
  theme_bw()

annual_plot_orig

# Create visualization of seasonal budgets
seasonal_plot_orig <- ggplot(seasonal_budget_orig, aes(x = as.factor(year), y = FCH4_F)) +
  geom_bar(stat = "identity", fill = "salmon") +
  facet_wrap(~season)+
  labs(title = "Annual CH4 Budgets - Original RF",
       x = "Season",
       y = expression(paste("CH"[4], " (g ", m^-2, " ", season^-1, ")"))) +
  theme_bw()

seasonal_plot_orig




# Display all results and plots together 
print(annual_budget_orig)

print(seasonal_budget_orig)

annual_plot_orig
seasonal_plot_orig
```
#plot all together for ease of comparison 
```{r}
library(dplyr)

#make sure date is in POSIXct
df_avg_orig$date <- as.POSIXct(df_avg_orig$date, format = "%Y-%m-%d")

class(df_avg_orig$date) #should report POSIXct


# Merge by date, adding just FCH4_F from df_avg_orig
df_avg2 <- df_avg %>%
 left_join(select(df_avg_orig, date, rfch4_orig = FCH4_F), by = "date")



# Calculate annual budgets
annual_budget_all <- df_avg2 %>%
  filter(year != 2023) %>%
  group_by(year) %>%
  summarise(
    XGBoost_total = sum(fch4_gapfilled) * 60 * 60 * 24 * (1/1000000000) * 12,
    XGBoost_interact_total = sum(fch4_gapfilled_interact) * 60 * 60 * 24 * (1/1000000000) * 12,
    RandomForest_total = sum(rfch4_gf) * 60 * 60 * 24 * (1/1000000000) * 12,
    Orig_RandomForest_total = sum(rfch4_orig) * 60 * 60 * 24 * (1/1000000000) * 12,
    n_days = n()
  )

# Calculate seasonal budgets
seasonal_budget_all <- df_avg2 %>%
  filter(year != 2023) %>%
  group_by(year, season) %>%
  summarise(
    XGBoost_total = sum(fch4_gapfilled) * 60 * 60 * 24 * (1/1000000000) * 12,
    XGBoost_interact_total = sum(fch4_gapfilled_interact) * 60 * 60 * 24 * (1/1000000000) * 12,
    RandomForest_total = sum(rfch4_gf) * 60 * 60 * 24 * (1/1000000000) * 12,
    Orig_RandomForest_total = sum(rfch4_orig) * 60 * 60 * 24 * (1/1000000000) * 12,
    n_days = n()
  )


# Create visualization of annual budgets
annual_plot_all <- ggplot(annual_budget_all %>%
                     pivot_longer(cols = contains("total"),
                                names_to = "Model",
                                values_to = "CH4_flux")) +
  geom_bar(aes(x = as.factor(year), y = CH4_flux, fill = Model),
           position = "dodge", stat = "identity") +
  labs(title = "Annual CH4 Budgets by Model",
       x = "Year",
       y = expression(paste("CH"[4], " (g ", m^-2, " ", yr^-1, ")"))) +
  theme_bw()

# Create visualization of seasonal budgets
seasonal_plot_all <- ggplot(seasonal_budget_all %>%
                       pivot_longer(cols = contains("total"),
                                  names_to = "Model",
                                  values_to = "CH4_flux")) +
  geom_bar(aes(x = as.factor(year), y = CH4_flux, fill = Model),
           position = "dodge", stat = "identity") +
  facet_wrap(~season) +
  labs(title = "Seasonal CH4 Budgets by Model",
       x = "Year",
       y = expression(paste("CH"[4], " (g ", m^-2, " ", season^-1, ")"))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# Display results and plots

print(annual_budget_all)
print(seasonal_budget_all)

annual_plot_all
seasonal_plot_all




```


#gapfilling with orig rf model 
```{r}

#plot half-hourly FCH4 data with the gap-filled data to compare 
ggplot(data = df_avg2)+
  geom_hline(yintercept = 0)+
  geom_point(aes(date, rfch4_orig,col='RF Gapfilled'))+
  geom_point(aes(date,fch4,col='Original'))+
    labs(
        y = expression(CH[4]~Flux~(nmolCH[4]~m^-2~y^-1)),
        x = "Time") +
  scale_color_manual(values=c('black','red'))+
  theme_bw()+
  #geom_vline(xintercept = as.POSIXct("2023-07-05"))+
  labs(title = "Orig RF Methane Flux (no interaction): Measured and Gap-filled Values")
```


















#Copied over for full budgets, but skipping CO2 budgets for now to focus on methane 

####Annual budgets
#### Net C Budgets: Annual bar plot
```{r}
#Creates numeric timestamp with the unit of minutes to integrate over half-hourly data

#this references the half hourly increments and takes the diff (so 30 min increments)
df_2023 <- df_2023 %>%
  mutate(time_minutes = as.numeric(difftime(TIMESTAMP, min(TIMESTAMP), units = "mins")))
df_2023$TIMESTAMP

# Units start as micro-moles of CO2/(m^2/s), converted to Grams of C/m^2/minute, integrated over minutes to get Grams of C/m^2

#creating new column of net C budget in order to make the annual budget bar plots
#net CO2 
df_2023 <- df_2023 %>%
  mutate(FC_F_no_NAs = ifelse(is.na(FC_F), 0, FC_F * 60 * (1/1000000) * 12))

#Net CO2 flux in g/m2
net_CO2 <- trapz(df_2023$time_minutes, df_2023$FC_F_no_NAs)

# Units start as nano-moles of CH4/(m^2/s), converted to Grams of C/m^2/minute, integrated over minutes to get Grams of C/m^2

#Net CH4 Flux
df_2023 <- df_2023 %>%
  mutate(FCH4_F_no_NAs = ifelse(is.na(FCH4_F), 0, FCH4_F*60*(1/1000000000)*12))


net_CH4 <- trapz(df_2023$time_minutes, df_2023$FCH4_F_no_NAs)

# Used IPCC Sixth Assessment Report (AR6) global warming potentials, 100 year time period - could use paper gwp* or delta equation for future analysis

net_CH4_CO2e <- net_CH4*27.2
sum = net_CO2+net_CH4


#Dataframe created to generate bar graph
net_wp_2023 <- data.frame(
  Category = c("CO2", "CH4", "Total"),
  Value = c(round(net_CO2, 2), round(net_CH4, 2), round(sum, 2) 
))

#Plot here

annual_Cbudget_2023 <- ggplot(net_wp_2023, aes(x = Category, y = Value)) +
  theme_bw()+
  geom_bar(stat = "identity", position = "dodge", fill = "turquoise3") +
  geom_hline(yintercept=0, colour = "black")+
  labs(
    x = "",
    #y = expression(FCH4~(gCH[4]-C/m^2/s)),
    y = expression(Net~Carbon~Flux~(gC~m^-2~y^-1))) +
  geom_label(aes(label = Value), vjust = ifelse(net_wp_2023$Value >= 0, -0.7, 1.5), colour = "black", fill ="white")+
  scale_y_continuous(
     #breaks = seq(-5, 5, 1.5),  # changing for 2023 
   # limits = c(-2, 5))+ #changing for year 2023
      breaks = seq(-80, 20, 10),  # Set limits for the primary axis - other years go up to -73
    limits = c(-80, 50))+ 
  labs(title = "Cumulative Annual Carbon Budget for 2023") +#for 2023
#labs(title = "2023")
   theme(
    axis.title.x = element_text(size = 16, face = "bold"),  # Make x-axis title larger and bold
    axis.title.y = element_text(size = 16, face = "bold"),  # Make y-axis title larger and bold
    axis.text.x = element_text(size = 14, face = "bold"),   # Make x-axis text larger and bold
    axis.text.y = element_text(size = 14, face = "bold")    # Make y-axis text larger and bold
  )# +
  
#coord_fixed(ratio = 0.6) #for making small 2023 fig 

annual_Cbudget_2023

```





