---
title: "Adjusting_FC_night" #Adjusting FC_night from original cutoff of SW-IN = 0 to SW_IN < 20 
output: html_document
date: "2026-04-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)
library(car)

Sys.setenv(TZ='UTC')
```

#Load data

```{r}

# #half-hourly dataframe --> gapfilled - corrected seasons, winter-yr
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.3.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


#daily avg dataframe --> gapfilled - updated seasons, sWC-nf, winter-year
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.5.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


```

#These versions of the dataset already have timestamp in POSIXct and already have a date variable, but leaving here for reference 

#Create useable timestamp variable 
```{r}
# df$TIMESTAMP_END = as.character(df$TIMESTAMP_END)
# df$TIMESTAMP_START = as.character(df$TIMESTAMP_START)
# 
# df$TIMESTAMP_END = as.POSIXct(df$TIMESTAMP_END, tz="UTC", format = "%Y%m%d%H%M")
# df$TIMESTAMP_START = as.POSIXct(df$TIMESTAMP_START, tz="UTC", format = "%Y%m%d%H%M")
# 

```



#reduce datasets to just time/day, FC, FC_F, and SW_IN
```{r}
# Reduced half-hourly dataset
df_reduced <- df %>%
  select(TIMESTAMP_END, FC, FC_F, SW_IN)


df_reduced$date = as.Date(df$TIMESTAMP_END)


```


#Create FC_night in the half-hourly cleaned Ameriflux-prepped dataset from step 6
```{r}
# Filter CO2 Flux data by incoming shortwave values to get nighttime fluxes

#Add to orig df
df <- df %>%
  mutate(FC_night20 = ifelse(SW_IN <= 20, FC, NA)) #variable for not gapfilled

df <- df %>%
  mutate(FC_night20_F = ifelse(SW_IN <= 20, FC_F, NA)) #variable for gapfilled 

#=============================================

#Add to reduced df for easier averaging and merging with orig df_avg 
df_reduced <- df_reduced %>%
  mutate(FC_night20 = ifelse(SW_IN <= 20, FC, NA)) #variable for not gapfilled

df_reduced <- df_reduced %>%
  mutate(FC_night20_F = ifelse(SW_IN <= 20, FC_F, NA)) #variable for gapfilled 
  

```



#Average by day to create continuous daily averages - NOTE: need to make sure to use "date" for openAir and timeAverage code to work*
```{r}
library(openair)
#need at least 10% data to create avg for FC_night 

#average dataframe with lighter threshold to get better nighttime data coverage
df_avg_night <- as.data.frame(date)
df_avg_night <-timeAverage(df_reduced, avg.time = "day", data.thresh = 10)

#Make sure date classes match 
class(df_avg$date)
class(df_avg_night$date)

#merge with orig df_avg; change date class to match orig df_avg 
df_avg <- df_avg %>%
  left_join(df_avg_night %>% 
              mutate(date = as.IDate(date)) %>%
              select(date, FC_night20, FC_night20_F),
            by = "date")


```


#check out the averages to make sure they look reasonable / screen 

```{r}

plot(df$date, df$FC_night20_F)

plot(df_avg$date, df_avg$FC_night20_F)

plot(df_avg$TS_3_1_1, df_avg$FC_night_F)
plot(df_avg$TA_ERA5, df_avg$FC_night_F)

```

# Save Data

```{r}
#these dataframes now include FC_night and FC_night_F 

#half-hourly df with night fluxes -- added ".2" to indicated from updated SWC
write.csv(x = df,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.4.csv',quote = F,row.names = F)

#daily avg df with night fluxes -- added ".2" to indicated from updated SWC
write.csv(x = df_avg,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.6.csv',quote = F,row.names = F)


```




