---
title: "Council_Cal_Year_vs_Flux_Year_means" #Calendar year budgets & means; "Flux Year" (winter year) budgets and means 
output: html_document
date: "2026-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)

#install.packages("writexl")
library(writexl)


Sys.setenv(TZ='UTC')
```

# Set working directory and then Load data


```{r}
#with gapfilled soil and air temp columns; SWC 3
# #half-hourly, gf (corrected seasons, Nov'25)
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


#daily avg dataframe --> gapfilled - updated seasons Nov 2025; added winter year Dec 2025
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.5.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


```


#====Calendar Year=======================

#Table: seasonal and annual total C budget & met variable tables 
```{r}
# Load necessary packages
library(dplyr)
library(lubridate)

#Annual and seasonal summaries with SE 

#no SE for CO2 and CH4 fluxes as these are sums **** then we calc averages with SE for the met variables 

# Annual summaries with standard errors
# Annual summaries with total C budget
annual_budgets <- df_avg %>%
  group_by(year = year(date)) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete year data 
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TS_ERA5 = mean(TS_ERA5, na.rm = TRUE),
    se_TS_ERA5 = sd(TS_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TS_ERA5))),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    se_TA_ERA5 = sd(TA_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TA_ERA5))),
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
   se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled))))
      )

print(annual_budgets)

# Seasonal summaries with total C budget
seasonal_budgets <- df_avg %>%
  group_by(year = year(date), season) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete year data 
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TS_ERA5 = mean(TS_ERA5, na.rm = TRUE),
    se_TS_ERA5 = sd(TS_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TS_ERA5))),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    se_TA_ERA5 = sd(TA_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TA_ERA5))),
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
    se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled)))),
    days = n()
  )


print(seasonal_budgets)


```
# cal year seasonal pub table 
```{r}
# Create a new table with combined mean ± SE format
cal_seasonal_budgets_formatted <- seasonal_budgets %>%
  mutate(
    soil_temp = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    soil_moisture = sprintf("%.2f ± %.2f", mean_soil_moisture, se_soil_moisture),
    air_temp = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    TS_ERA5 = sprintf("%.2f ± %.2f", mean_TS_ERA5, se_TS_ERA5),
    TA_ERA5 = sprintf("%.2f ± %.2f", mean_TA_ERA5, se_TA_ERA5),
    TA_GF = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    TS3_GF = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
  ) %>%
  dplyr::select(year, season, CO2_budget, CH4_budget, total_C_budget, soil_temp, soil_moisture, air_temp, TS_ERA5, TA_ERA5, TA_GF, TS3_GF, days)

cal_seasonal_budgets_formatted
```
#save table 
```{r}
# Save to Excel
write_xlsx(cal_seasonal_budgets_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/cal_yr_seasonal_budgets_formatted.xlsx")
```

#cal year annual pub table 
```{r}
#Annual budgets 
# Create a new table with combined mean ± SE format
cal_annual_budgets_formatted <- annual_budgets %>%
  mutate(
    soil_temp = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    soil_moisture = sprintf("%.2f ± %.2f", mean_soil_moisture, se_soil_moisture),
    air_temp = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    TS_ERA5 = sprintf("%.2f ± %.2f", mean_TS_ERA5, se_TS_ERA5),
    TA_ERA5 = sprintf("%.2f ± %.2f", mean_TA_ERA5, se_TA_ERA5),
    TA_GF = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    TS3_GF = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
  ) %>%
  dplyr::select(year, CO2_budget, CH4_budget, total_C_budget, soil_temp, soil_moisture, air_temp, TS_ERA5, TA_ERA5, TA_GF, TS3_GF)

cal_annual_budgets_formatted
```

```{r}
# Save to Excel
write_xlsx(annual_budgets_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/cal_yr_annual_budgets_formatted2.xlsx")

```

# % gapfilled vs observed for TS3 and TA

```{r}
# Calculate percentages for Air Temperature (TA)
#HH - year, season 
TA_summaryHH <- df %>%
   filter( year != 2023) %>%
  group_by(year, season) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TA_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TA_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TA_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Air_Temperature")

#daily avg - year, season
TA_summary <- df_avg %>%
   filter( year != 2023) %>%
  group_by(year, season) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TA_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TA_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TA_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Air_Temperature")
TA_summary

#HH - year only 
TA_summaryHH_annual <- df %>%
   filter( year != 2023) %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TA_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TA_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TA_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Air_Temperature")

#daily avg - year only
TA_summary_annual <- df_avg %>%
   filter( year != 2023) %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TA_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TA_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TA_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Air_Temperature")


# Calculate percentages for Soil Temperature (TS_3_1_1)

#HH - year, season 
TS_summaryHH <- df %>%
  filter( year != 2023) %>%
  group_by(year, season) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TS_3_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TS_3_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TS_3_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Soil_Temperature")


#daily avg - year, season 
TS_summary <- df_avg %>%
   filter( year != 2023) %>%
  group_by(year, season ) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TS_3_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TS_3_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TS_3_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Soil_Temperature")

#HH - year only 
TS_summaryHH_annual <- df %>%
  filter( year != 2023) %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TS_3_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TS_3_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TS_3_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Soil_Temperature")


#daily avg - year only 
TS_summary_annual <- df_avg %>%
   filter( year != 2023) %>%
  group_by(year, season ) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TS_3_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TS_3_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TS_3_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Soil_Temperature")

```

#Add % gapfilled (from HH datasets) to seasonal and annual averages table - all SWC

```{r}
# Join the season, year datasets, adding only the gapfilled_percent column 
cal_seasonal_budgets_with_gapfill <- seasonal_budgets %>%
  left_join(TA_summaryHH %>% select(year, season, TA_gapfill_percent = gapfilled_percent), 
            by = c("year", "season")) %>%
   left_join(TS_summaryHH %>% select(year, season, TS3_gapfill_percent = gapfilled_percent), 
            by = c("year", "season"))

# Check the result
print(cal_seasonal_budgets_with_gapfill)


# Join the annual datasets, adding only the gapfilled_percent column
cal_annual_budgets_with_gapfill <- seasonal_budgets %>%
  left_join(TA_summaryHH_annual %>% select(year, TA_gapfill_percent = gapfilled_percent), 
            by = c("year")) %>%
   left_join(TS_summaryHH_annual %>% select(year, TS3_gapfill_percent = gapfilled_percent), 
            by = c("year"))

# Check the result
print(cal_annual_budgets_with_gapfill)


# ~~~~~~~~~~~~~Add to formatted tables ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#Annual 
# Join the annual datasets, adding only the gapfilled_percent column
cal_annual_budgets_formatted2 <- cal_annual_budgets_formatted %>%
  left_join(TA_summaryHH_annual %>% select(year, TA_gapfill_percent = gapfilled_percent), 
            by = c("year")) %>%
   left_join(TS_summaryHH_annual %>% select(year, TS3_gapfill_percent = gapfilled_percent), 
             by = c("year"))

# Check the result
print(cal_annual_budgets_formatted2)


#Seasonal 

# Join the seasonal datasets, adding only the gapfilled_percent column
cal_seasonal_budgets_formatted2 <- cal_seasonal_budgets_formatted %>%
  left_join(TA_summaryHH %>% select(year, season, TA_gapfill_percent = gapfilled_percent), 
            by = c("year", "season")) %>%
   left_join(TS_summaryHH %>% select(year, season, TS3_gapfill_percent = gapfilled_percent), 
            by = c("year", "season"))

# Check the result
print(cal_seasonal_budgets_formatted2)

```



# Save to Excel
```{r}
#annual 
write_xlsx(cal_annual_budgets_formatted2, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/cal_year_annual_budgetsGF_formatted.xlsx")

#seasonal 
write_xlsx(cal_seasonal_budgets_formatted2, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/cal_year_seasonal_budgetsGF_formatted.xlsx")

```

#Winter adjustments based on proportions from 2018-2019 winter 

#Winter adj based on daily mean flux of growing season 

#FCH4 - daily mean winter adj 

#based on 2019 dailg avg growing season fluxes, winter daily avg CH4 flux can be ~44% of growing season daily avg flux (0.445) --> multiply winter daily avg fluxes by 0.445 & re-calc the daily avg CH4 flux to get an adjusted potential winter daily avg flux 

```{r}
# CH4 in nano-moles of CH4/m^2/s)
#Net CH4 Flux - gapfilled

# Calculate 0.445 (~44%) of daily average CH4 flux (FCH4_F) during growing season season with standard error

seasonal_FCH4_summary_gf_winteradjCH4 <- df_avg %>%
  filter(season == "Growing Season") %>%
  group_by(year, season) %>%
  summarize(
    mean_flux = mean(FCH4_F*0.445, na.rm = TRUE),
    se_flux = sd(FCH4_F*0.445, na.rm = TRUE)/sqrt(sum(!is.na(FCH4_F))),
    n_days = n()  # number of days per season
  )

#renaming the ~44% of growing season avg variable name 
seasonal_FCH4_summary_gf_winteradjCH4$season[seasonal_FCH4_summary_gf_winteradjCH4$season=="Growing Season"] <- "Winter Adj FCHF DailyAvg"


seasonal_FCH4_summary_gf_winteradjCH4 

# Format the results with mean ± SE
seasonal_FCH4_summary_gf_winteradjCH4_formatted <- seasonal_FCH4_summary_gf_winteradjCH4 %>%
  mutate(
    flux = sprintf("%.2f ± %.2f", mean_flux, se_flux)
  ) %>%
  select(season, flux)
         
seasonal_FCH4_summary_gf_winteradjCH4_formatted
```

#save table 
```{r}
# # Save to Excel
library(writexl)
write_xlsx(seasonal_FCH4_summary_gf_winteradjCH4_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteradj_dailyAvg_FCH4_summary_gf.xlsx")

```

# NEE - daily flux winter-adj

#based on 2019 dailg avg growing season fluxes, winter daily avg CO2 flux can be ~47% of growing season daily avg flux (0.473) --> multiply winter daily avg fluxes by 0.473 & re-calc the daily avg flux to get an adjusted potential winter daily avg flux 

```{r}
#FC in umol/m2/s

# Calculate 0.473 (~47%) of daily average daily NEE (FC_F) during growing season season with standard error

seasonal_NEE_summary_gf_winteradjNEE <- df_avg %>%
  filter(season == "Growing Season") %>%
  group_by(year, season) %>%
  summarize(
    mean_flux = abs(mean(FC_F*0.473, na.rm = TRUE)),
    se_flux = sd(FC_F*0.473, na.rm = TRUE)/sqrt(sum(!is.na(FC_F))),
    n_days = n()  # number of days per season
  )

#renaming the ~47% of growing season avg variable name 
seasonal_NEE_summary_gf_winteradjNEE$season[seasonal_NEE_summary_gf_winteradjNEE$season=="Growing Season"] <- "Winter Adj NEE DailyAvg"


seasonal_NEE_summary_gf_winteradjNEE 

# Format the results with mean ± SE
seasonal_NEE_summary_gf_winteradjNEE_formatted <- seasonal_NEE_summary_gf_winteradjNEE  %>%
  mutate(
    flux = sprintf("%.2f ± %.2f", mean_flux, se_flux)
  ) %>%
  select(season, flux)
         
seasonal_NEE_summary_gf_winteradjNEE_formatted
```

#save table 
```{r}
# # Save to Excel
# library(writexl)
write_xlsx(seasonal_NEE_summary_gf_winteradjNEE_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteradj_dailyAvg_NEE_summary_gf.xlsx")

```


















#=========================

#Cal year 

```{r}
# Seasonal Trends Faceted by Season - All Variables

library(ggplot2)
library(dplyr)

# Prepare seasonal summaries for all variables (2017-2022)
seasonal_summary_all <- df_avg %>%
 # filter(winter_year >= 2017 & winter_year <= 2022) %>%
group_by(winter_year, season) %>%
  summarise(
    # ERA5 GF Air temperature (gap-filled)
    TA_gapfilled_mean = mean(TA_gapfilled, na.rm = TRUE),
    TA_gapfilled_se = sd(TA_gapfilled, na.rm = TRUE) / sqrt(sum(!is.na(TA_gapfilled))),
    
    # measured Air temperature 
    TA_mean = mean(TA, na.rm = TRUE),
    TA_se = sd(TA, na.rm = TRUE) / sqrt(sum(!is.na(TA))),
    
    # ERA5 GF Soil temperature (gap-filled only)
    TS_3_gapfilled_mean = mean(TS_3_gapfilled, na.rm = TRUE),
    TS_3_gapfilled_se = sd(TS_3_gapfilled, na.rm = TRUE) / sqrt(sum(!is.na(TS_3_gapfilled))),
    
    # Measured Soil temperature 
    TS_3_mean = mean(TS_3_1_1, na.rm = TRUE),
    TS_3_se = sd(TS_3_1_1, na.rm = TRUE) / sqrt(sum(!is.na(TS_3_1_1))),
    
    # Soil moisture
    # SWC_3_mean = mean(SWC_3_1_1, na.rm = TRUE),
    # SWC_3_se = sd(SWC_3_1_1, na.rm = TRUE) / sqrt(sum(!is.na(SWC_3_1_1))),
    
    # Soil moisture - unfrozen, excluding winter and when soil temp is <= 2 deg
    SWC_3_nf_mean = mean(SWC_3_nf, na.rm = TRUE),
    SWC_3_nf_se = sd(SWC_3_nf, na.rm = TRUE) / sqrt(sum(!is.na(SWC_3_nf))),
    
    # VPD
    VPD_mean = mean(VPD, na.rm = TRUE),
    VPD_se = sd(VPD, na.rm = TRUE) / sqrt(sum(!is.na(VPD))),
    
    # RH - relative humidity 
    RH_mean = mean(RH, na.rm = TRUE),
    RH_se = sd(RH, na.rm = TRUE) / sqrt(sum(!is.na(RH))),
    
    # Sensible heat
    H_mean = mean(H, na.rm = TRUE),
    H_se = sd(H, na.rm = TRUE) / sqrt(sum(!is.na(H))),
    
    # Latent heat
    LE_mean = mean(LE, na.rm = TRUE),
    LE_se = sd(LE, na.rm = TRUE) / sqrt(sum(!is.na(LE))),
    
    # Ground heat
    G_mean = mean(G_1_1_1, na.rm = TRUE),
    G_se = sd(G_1_1_1, na.rm = TRUE) / sqrt(sum(!is.na(G_1_1_1))),
    
    
    
    .groups = 'drop'
  )

# Define consistent year colors
winteryear_colors <- c("2017" = "#E41A1C",  # Red
                 "2018" = "#377EB8",  # Blue
                 "2019" = "#4DAF4A",  # Green
                 "2020" = "#984EA3",  # Purple
                 "2021" = "#FF7F00",  # Orange
                 "2022" = "#A65628")  # Brown

# Convert year to factor for consistent color mapping
seasonal_summary_all$winteryear_factor <- as.factor(seasonal_summary_all$winter_year)
```



# %gapfill of other met variables 
```{r}
gapfill_summary <- df_avg %>%
  # Just group directly by winter_year (which already exists!)
  group_by(winter_year, season) %>%
  summarise(
    # Total days 
    total_days = n(),
    
    
    #Air temp
    VPD_measured = sum(!is.na(VPD)),
    VPD_missing = sum(is.na(VPD)),
    VPD_pct_missing = round((VPD_missing / total_days) * 100, 1),
    VPD_pct_coverage = round((VPD_measured / total_days) * 100, 1),
    
    #Soil temp
    VPD_measured = sum(!is.na(VPD)),
    VPD_missing = sum(is.na(VPD)),
    VPD_pct_missing = round((VPD_missing / total_days) * 100, 1),
    VPD_pct_coverage = round((VPD_measured / total_days) * 100, 1),
    
    
  #SWC_nf (SWC above 2degC)
    SWC_nf_measured = sum(!is.na(SWC_3_nf)),
    SWC_nf_missing = sum(is.na(SWC_3_nf)),
    SWC_nf_pct_missing = round((SWC_nf_missing / total_days) * 100, 1),
    SWC_nf_pct_coverage = round((SWC_nf_measured / total_days) * 100, 1),
  
    #VPD
    VPD_measured = sum(!is.na(VPD)),
    VPD_missing = sum(is.na(VPD)),
    VPD_pct_missing = round((VPD_missing / total_days) * 100, 1),
    VPD_pct_coverage = round((VPD_measured / total_days) * 100, 1),
  
      #RH
    RH_measured = sum(!is.na(RH)),
    RH_missing = sum(is.na(RH)),
    RH_pct_missing = round((RH_missing / total_days) * 100, 1),
    RH_pct_coverage = round((RH_measured / total_days) * 100, 1),
  
     #H
    H_measured = sum(!is.na(H)),
    H_missing = sum(is.na(H)),
    H_pct_missing = round((H_missing / total_days) * 100, 1),
    H_pct_coverage = round((H_measured / total_days) * 100, 1),
  
     #LE
    LE_measured = sum(!is.na(LE)),
    LE_missing = sum(is.na(LE)),
    LE_pct_missing = round((LE_missing / total_days) * 100, 1),
    LE_pct_coverage = round((LE_measured / total_days) * 100, 1),
  
      #G
    G_1_1_1_measured = sum(!is.na(G_1_1_1)),
    G_1_1_1_missing = sum(is.na(G_1_1_1)),
    G_1_1_1_pct_missing = round((G_1_1_1_missing / total_days) * 100, 1),
    G_1_1_1_pct_coverage = round((G_1_1_1_measured / total_days) * 100, 1),
    
    
    
    .groups = "drop"
  ) 


# Display summary
print(gapfill_summary)

#only % coverage / % missing 
gapfill_summary2 <- gapfill_summary %>%
  select(
    winter_year,
         season,
         total_days,
        # ends_with("pct_missing"), 
         ends_with("pct_coverage")
  )

gapfill_summary2

```

```{r}
#full summary  
write_xlsx(gapfill_summary, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteryear_met_vars_data_coverage.xlsx")

#full summary  
write_xlsx(gapfill_summary2, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteryear_met_vars_gapfill_pct.xlsx")

```


#Make formatted table of seasonal means and % coverage of other met variables 
```{r}
library(tidyverse)
library(knitr)
library(kableExtra)

# ============================================================================
# COMBINE SEASONAL MEANS WITH DATA COVERAGE
# ============================================================================

# Join the two datasets
combined_table <- seasonal_summary_all %>%
  select(-starts_with("TA"), -starts_with("TS_3")) %>%
  filter(winter_year != "2016") %>%
  left_join(gapfill_summary2, by = c("winter_year", "season")) %>%
  # Create formatted columns: "mean ± SE (coverage%)"
  mutate(

    # Soil Moisture (non-frozen)
    `SWC_3_nf` = sprintf("%.2f ± %.2f (%.1f%%)", 
                         SWC_3_nf_mean, SWC_3_nf_se, SWC_nf_pct_coverage),
    
    # VPD
    `VPD` = sprintf("%.2f ± %.2f (%.1f%%)", 
                    VPD_mean, VPD_se, VPD_pct_coverage),
    
    # Relative Humidity
    `RH` = sprintf("%.2f ± %.2f (%.1f%%)", 
                   RH_mean, RH_se, RH_pct_coverage),
    
    # Sensible Heat
    `H` = sprintf("%.2f ± %.2f (%.1f%%)", 
                  H_mean, H_se, H_pct_coverage),
    
    # Latent Heat
    `LE` = sprintf("%.2f ± %.2f (%.1f%%)", 
                   LE_mean, LE_se, LE_pct_coverage),
    
    # Ground Heat Flux
    `G` = sprintf("%.2f ± %.2f (%.1f%%)", 
                  G_mean, G_se, G_1_1_1_pct_coverage)
  ) %>%
  # Select only the formatted columns
  select(winter_year, season, total_days,
                  `SWC_3_nf`, `VPD`, `RH`, `H`, `LE`, `G`)

# View the table
print(combined_table)


# Rename columns for publication
# pub_table <- combined_table %>%
#   rename(
#     `Year` = winter_year,
#     `Season` = season,
#     `Days` = total_days,
#     `Air Temp GF (°C)` = TA_gapfilled,
#     `Air Temp Meas (°C)` = TA,
#     `Soil Temp GF (°C)` = TS_3_gapfilled,
#     `Soil Temp Meas (°C)` = TS_3,
#     `Soil Moisture (%)` = SWC_3_nf,
#     `VPD (hPa)` = VPD,
#     `RH (%)` = RH,
#     `H (W/m²)` = H,
#     `LE (W/m²)` = LE,
#     `G (W/m²)` = G
#   )


```

#save table of mean met vars with % data coverage
```{r}
# Save to Excel
write_xlsx(combined_table, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteryear_met_var_means_pct_coverage_formatted.xlsx")
```

# ===========================================

# "Flux Year" (winter year)


#Table: seasonal and annual total C budget & met variable tables 
```{r}
# Load necessary packages
library(dplyr)
library(lubridate)

#Annual and seasonal summaries with SE 

#no SE for CO2 and CH4 fluxes as these are sums **** then we calc averages with SE for the met variables 

# Annual summaries with standard errors
# Annual summaries with total C budget
WY_annual_budgets <- df_avg %>%
  group_by(year = winter_year) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete year data 
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TS_ERA5 = mean(TS_ERA5, na.rm = TRUE),
    se_TS_ERA5 = sd(TS_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TS_ERA5))),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    se_TA_ERA5 = sd(TA_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TA_ERA5))),
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
   se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled))))
      )

print(WY_annual_budgets)

# Seasonal summaries with total C budget
WY_seasonal_budgets <- df_avg %>%
  group_by(year = winter_year, season) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete year data 
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    total_C_budget = CO2_budget + CH4_budget,  # Adding total C budget
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TS_ERA5 = mean(TS_ERA5, na.rm = TRUE),
    se_TS_ERA5 = sd(TS_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TS_ERA5))),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    se_TA_ERA5 = sd(TA_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TA_ERA5))),
    mean_TA_gapfilled   = mean(TA_gapfilled, na.rm = TRUE),  
    se_TA_gapfilled   = sd(TA_gapfilled, na.rm = TRUE)/sqrt(sum(!is.na(TA_gapfilled))), 
    mean_TS_3_gapfilled = mean(TS_3_gapfilled, na.rm = TRUE), 
    se_TS_3_gapfilled = sd(TS_3_gapfilled, na.rm = TRUE /sqrt(sum(!is.na(TS_3_gapfilled)))),
    days = n()
  )


print(WY_seasonal_budgets)


```


# cal year seasonal pub table 
```{r}
# Create a new table with combined mean ± SE format
WY_seasonal_budgets_formatted <- WY_seasonal_budgets %>%
  mutate(
    soil_temp = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    soil_moisture = sprintf("%.2f ± %.2f", mean_soil_moisture, se_soil_moisture),
    air_temp = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    TS_ERA5 = sprintf("%.2f ± %.2f", mean_TS_ERA5, se_TS_ERA5),
    TA_ERA5 = sprintf("%.2f ± %.2f", mean_TA_ERA5, se_TA_ERA5),
    TA_GF = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    TS3_GF = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
  ) %>%
  dplyr::select(year, season, CO2_budget, CH4_budget, total_C_budget, soil_temp, soil_moisture, air_temp, TS_ERA5, TA_ERA5, TA_GF, TS3_GF, days)

WY_seasonal_budgets_formatted

```


#save table 
```{r}
# Save to Excel
write_xlsx(WY_seasonal_budgets_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteryear_seasonal_budgets_formatted.xlsx")
```

#Winter year annual pub table 
```{r}
#Annual budgets 
# Create a new table with combined mean ± SE format
WY_annual_budgets_formatted <- WY_annual_budgets %>%
  mutate(
    soil_temp = sprintf("%.2f ± %.2f", mean_soil_temp, se_soil_temp),
    soil_moisture = sprintf("%.2f ± %.2f", mean_soil_moisture, se_soil_moisture),
    air_temp = sprintf("%.2f ± %.2f", mean_TA, se_TA),
    TS_ERA5 = sprintf("%.2f ± %.2f", mean_TS_ERA5, se_TS_ERA5),
    TA_ERA5 = sprintf("%.2f ± %.2f", mean_TA_ERA5, se_TA_ERA5),
    TA_GF = sprintf("%.2f ± %.2f", mean_TA_gapfilled, se_TA_gapfilled),
    TS3_GF = sprintf("%.2f ± %.2f", mean_TS_3_gapfilled, se_TS_3_gapfilled),
  ) %>%
  dplyr::select(year, CO2_budget, CH4_budget, total_C_budget, soil_temp, soil_moisture, air_temp, TS_ERA5, TA_ERA5, TA_GF, TS3_GF)

WY_annual_budgets_formatted 
```

```{r}
# Save to Excel
write_xlsx(WY_annual_budgets_formatted, "C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/winteryear_annual_budgets_formatted.xlsx")

```
