---
title: "Council_EC Flux_Analysis.2_Github" #cleaning up the code for analysis while retaining the old one 
output: html_document
date: "2025-02-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Notes on site: Wind direction primarily blows from NW, except for Dec - Mar, when it primarily blows from SE direction. The site is largely tussock tundra but to the south there is thermokarst 


## based on Dani's Churchill CO2 Analysis code on github 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# add root directory here for github
```

Notes:

```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)


Sys.setenv(TZ='UTC')
```

# Set working directory and then Load data

```{r}
#using the ".2" version to reflect the updated SWC used in RF 

#original half-hourly dataframe 
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

#needs continuous day variable - daily avg, calculated in processing steps
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

# #monthly avg, calculated in processing steps 
# df_monthly_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_monthly_AVG_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

#original half-hourly dataframe using the SWC1
#dforig = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))
```


```{r}
#quick look at the different temp profiles (all at 15 cm depth, diff locations)
sum(is.na(df$TS_1_1_1)) #60277 NA's in temp by margin pond 
sum(is.na(df$TS_2_1_1)) #48729 NA's in temp  by lichen/berries  
sum(is.na(df$TS_3_1_1)) #50445 NA's in temp by tussock 
sum(is.na(df$TS_4_1_1)) #52796 NA's in temp by foot of EC tower
sum(is.na(df$TS_5_1_1)) #new temp in latest upload with 2023 data, don't have info on what it is yet** 


#SWC also broken up by location
#  SWC_1_1_1 % Soil water content (15cm depth) – margin pond
# SWC_2_1_1 % Soil water content (15cm depth) – lichen/berries
# SWC_3_1_1 % Soil water content (15cm depth) - tussock


```

```{r}
#checking timestamp 
ggplot(data = df,aes(TIMESTAMP_END,FC_F))+
  geom_point()+
  scale_x_datetime(limits = as.POSIXct(c('2021-09-28','2021-10-05')))+
  scale_y_continuous(limits = c(-12,5))
```

#create a useable timestamp - not necessary here as it's already in the correct format from our BASE processing code 
```{r}
#ignore this if df is already in correct format, as this might mess up the formatting here and cause issues with analysis
# df$TIMESTAMP_END = as.character(df$TIMESTAMP_END)
# df$TIMESTAMP_START = as.character(df$TIMESTAMP_START)
# 
# df$TIMESTAMP_END = as.POSIXct(df$TIMESTAMP_END, tz="UTC", format = "%Y%m%d%H%M")
# df$TIMESTAMP_START = as.POSIXct(df$TIMESTAMP_START, tz="UTC", format = "%Y%m%d%H%M")
```

#Add Seasonal delineations to df and df_avg (see older version of code for work on delineations)
```{r}
# add year column if not already present
df <- df %>%
  mutate(year = format(date, "%Y"))

# Create DOY for df (half-hourly data)
df$DOY <- yday(df$date)


# Create seasonal assignments using both year and DOY
df<- df %>%
  mutate(
    season = case_when(
      # 2017 seasons
      year == "2017" & (DOY >= 258 & DOY <= 290) ~ 'Fall Senescence',
      year == "2017" & (DOY >= 131 & DOY <= 257) ~ 'Growing Season',
      year == "2017" & (DOY >= 291 | DOY <= 130) ~ 'Winter',
      
      
      # 2018 seasons
      year == "2018" &  (DOY >= 257 & DOY <= 300) ~ 'Fall Senescence',
      year == "2018" & (DOY >= 152 & DOY <= 256) ~ 'Growing Season',
      year == "2018" & (DOY >= 301 | DOY <= 151) ~ 'Winter',
      
      # 2019 seasons
      year == "2019" & (DOY >= 246 & DOY <= 284) ~ 'Fall Senescence',
      year == "2019" & (DOY >= 145 & DOY <= 245) ~ 'Growing Season',
      year == "2019" & (DOY >= 285 | DOY <= 144) ~ 'Winter',
      
      # 2020 seasons
      year == "2020" & (DOY >= 252 & DOY <= 299) ~ 'Fall Senescence',
      year == "2020" & (DOY >= 136 & DOY <= 251) ~ 'Growing Season',
      year == "2020" & (DOY >= 300 | DOY <= 135) ~ 'Winter',
      
      # 2021 seasons
      year == "2021" & (DOY >= 260 & DOY <= 283) ~ 'Fall Senescence',
      year == "2021" &(DOY >= 142 & DOY <= 259) ~ 'Growing Season',
      year == "2021" &(DOY >= 284 | DOY <= 141) ~ 'Winter',
      
      # 2022 seasons
      year == "2022" & (DOY >= 245 & DOY <= 281) ~ 'Fall Senescence',
      year == "2022" & (DOY >= 144 & DOY <= 244) ~ 'Growing Season',
      year == "2022" & (DOY >= 282 | DOY <= 143) ~ 'Winter',
      
      # not doing one for 2023 ** incomplete year 
      TRUE ~ NA_character_
    )
  )

####################easona to daily avg dataset###########################

# add year column if not already present
df_avg <- df_avg %>%
  mutate(year = format(date, "%Y"))


# Create DOY For df_avg (daily data)
df_avg$DOY <- yday(df_avg$date)


# Create seasonal assignments using both year and DOY
df_avg <- df_avg %>%
  mutate(
    season = case_when(
      # 2017 seasons
      year == "2017" & (DOY >= 258 & DOY <= 290) ~ 'Fall Senescence',
      year == "2017" & (DOY >= 131 & DOY <= 257) ~ 'Growing Season',
      year == "2017" & (DOY >= 291 | DOY <= 130) ~ 'Winter',
      
      
      # 2018 seasons
      year == "2018" &  (DOY >= 257 & DOY <= 300) ~ 'Fall Senescence',
      year == "2018" & (DOY >= 152 & DOY <= 256) ~ 'Growing Season',
      year == "2018" & (DOY >= 301 | DOY <= 151) ~ 'Winter',
      
      # 2019 seasons
      year == "2019" & (DOY >= 246 & DOY <= 284) ~ 'Fall Senescence',
      year == "2019" & (DOY >= 145 & DOY <= 245) ~ 'Growing Season',
      year == "2019" & (DOY >= 285 | DOY <= 144) ~ 'Winter',
      
      # 2020 seasons
      year == "2020" & (DOY >= 252 & DOY <= 299) ~ 'Fall Senescence',
      year == "2020" & (DOY >= 136 & DOY <= 251) ~ 'Growing Season',
      year == "2020" & (DOY >= 300 | DOY <= 135) ~ 'Winter',
      
      # 2021 seasons
      year == "2021" & (DOY >= 260 & DOY <= 283) ~ 'Fall Senescence',
      year == "2021" &(DOY >= 142 & DOY <= 259) ~ 'Growing Season',
      year == "2021" &(DOY >= 284 | DOY <= 141) ~ 'Winter',
      
      # 2022 seasons
      year == "2022" & (DOY >= 245 & DOY <= 281) ~ 'Fall Senescence',
      year == "2022" & (DOY >= 144 & DOY <= 244) ~ 'Growing Season',
      year == "2022" & (DOY >= 282 | DOY <= 143) ~ 'Winter',
      
      # not doing one for 2023 ** incomplete year 
      TRUE ~ NA_character_
    )
  )
```

#Create yearly dataframes

#create TIMESTAMP variable that is = to TIMESTAMP_END

```{r}
 
df$TIMESTAMP = df$TIMESTAMP_END

year_df <- function(df, year) {
  df %>%
    filter(format(date, "%Y") == as.character(year)) %>%
    mutate(DOY = yday(date))
}


#for orig dataset using SWC1

# year_df2 <- function(df, year) {
#   dforig %>%
#     filter(format(date, "%Y") == as.character(year)) %>%
#     mutate(DOY = yday(date))
# }
```

#Creates data frames of each year containing the daily averages of each value- useful for focusing on one year at a time
```{r}
#create dataframes for each year for the data you have so you can look at annual trends 

#map out the timestamp of the daily avg dataset
df_avg$date <- as.POSIXct(df_avg$date, format = "%Y-%m-%d") #tells R the order of the timestamp: yr-month-day 

df_avg_2017 <- year_df(df_avg, 2017)
df_avg_2018 <- year_df(df_avg, 2018)
df_avg_2019 <- year_df(df_avg, 2019)
df_avg_2020 <- year_df(df_avg, 2020)
df_avg_2021 <- year_df(df_avg, 2021)
df_avg_2022 <- year_df(df_avg, 2022)
df_avg_2023 <- year_df(df_avg, 2023)


#map out the timestamp of the half-hourly dataset
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d %H:%M:%OS") #yr month day hour min sec
#dforig$date <- as.POSIXct(df$date, format = "%Y-%m-%d %H:%M:%OS") #yr month day hour min sec --> dataset using SWC_1

df_2017 <- year_df(df, 2017)
df_2018 <- year_df(df, 2018)
df_2019 <- year_df(df, 2019)
df_2020 <- year_df(df, 2020)
df_2021 <- year_df(df, 2021)
df_2022 <- year_df(df, 2022)
df_2023 <- year_df(df, 2023)


```

#Comparing SWC and soil temp from the different locations 
```{r}
#SWC 
ggplot(data = df) +
  geom_line(aes(TIMESTAMP, SWC_1_1_1, col = 'SWC1 margin pond'))+
  geom_line(aes(TIMESTAMP, SWC_2_1_1, col = 'SWC2 lichen berries')) +
  geom_line(aes(TIMESTAMP, SWC_3_1_1, col = 'SWC3 tussock')) +
  labs(title = "Comparing SWC locations")+
  labs( x = "Timestamp",
        y = "SWC")+
  scale_y_continuous()+
  theme_minimal()


#soil temp
ggplot(data = df) +
    geom_line(aes(TIMESTAMP, TS_5_1_1, col = 'TS5 unknown')) + #putting it at front so other colored lines are not obscured by TS5
  geom_line(aes(TIMESTAMP, TS_1_1_1, col = 'TS1 margin pond'))+
  geom_line(aes(TIMESTAMP, TS_2_1_1, col = 'TS2 lichen berries')) +
  geom_line(aes(TIMESTAMP, TS_3_1_1, col = 'TS3 tussock')) +
  geom_line(aes(TIMESTAMP, TS_4_1_1, col = 'TS4 EC tower')) +
  #geom_line(aes(TIMESTAMP, TS_5_1_1, col = 'TS5 unknown')) +
  labs(title = "Comparing Soil temp among locations")+
  labs( x = "Timestamp",
        y = "Soil temp (C)")+
  scale_y_continuous()+
  theme_minimal()


```

#Seasonal and annual summaries of CO2, CH4, TS_3, and SWC_3
```{r}
# Load necessary packages
library(dplyr)
library(ggplot2)
library(lubridate)

#annual summaries of CO2 and CH4 fluxes
annual_budgets <- df_avg %>%
  group_by(year = year(date)) %>%
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    mean_TS_ERA5 = mean (TS_ERA5, na.rm = TRUE),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    mean_TA = mean(TA, na.rm = TRUE),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE)
  )

#seasonal summaries
seasonal_budgets <- df_avg %>%
  group_by(year = year(date), season) %>%
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    mean_TS_ERA5 = mean (TS_ERA5, na.rm = TRUE),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    mean_TA = mean(TA, na.rm = TRUE),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    days = n()
  )

print(annual_budgets)
print(seasonal_budgets)
```


#Annual and seaosnal summaries with SE 
```{r}
#no SE for CO2 and CH4 fluxes as these are sums **** then we calc averages with SE for the met variables 

# Annual summaries with standard errors
annual_budgets <- df_avg %>%
  group_by(year = year(date)) %>%
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TS_ERA5 = mean(TS_ERA5, na.rm = TRUE),
    se_TS_ERA5 = sd(TS_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TS_ERA5))),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    se_TA_ERA5 = sd(TA_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TA_ERA5)))
  )

# Seasonal summaries with standard errors
seasonal_budgets <- df_avg %>%
  group_by(year = year(date), season) %>%
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    se_soil_temp = sd(TS_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(TS_3_1_1))),
    mean_TS_ERA5 = mean(TS_ERA5, na.rm = TRUE),
    se_TS_ERA5 = sd(TS_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TS_ERA5))),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    se_soil_moisture = sd(SWC_3_1_1, na.rm = TRUE)/sqrt(sum(!is.na(SWC_3_1_1))),
    mean_TA = mean(TA, na.rm = TRUE),
    se_TA = sd(TA, na.rm = TRUE)/sqrt(sum(!is.na(TA))),
    mean_TA_ERA5 = mean(TA_ERA5, na.rm = TRUE),
    se_TA_ERA5 = sd(TA_ERA5, na.rm = TRUE)/sqrt(sum(!is.na(TA_ERA5))),
    days = n()
  )

print(annual_budgets)
print(seasonal_budgets)
```


#***Annual budgets, seasonal budgets, figures, and correlations using daily avg df_avg
```{r}
# Load necessary packages
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)

# Calculate annual budgets
annual_budgets <- df_avg %>%
  group_by(year = year(date)) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete seasons 
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    mean_air_temp = mean(TA, na.rm = TRUE),
    mean_wind_speed = mean(WS, na.rm = TRUE),
    total_days = n()
  )

# Calculate seasonal budgets
seasonal_budgets <- df_avg %>%
  group_by(year = year(date), season) %>%
  filter(year != 2023) %>% #removing 2023 due to incomplete seasons 
  summarize(
    CO2_budget = sum(FC_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*60*24*(1/1000000000)*12), na.rm = TRUE),
    GPP_sum = sum(GPP_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    RECO_sum = sum(RECO_F * (60*60*24*(1/1000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    mean_air_temp = mean(TA, na.rm = TRUE),
    mean_wind_speed = mean(WS, na.rm = TRUE),
    days = n()
  )

# Create visualizations

# 1. Annual CO2 and CH4 budgets
ggplot(annual_budgets, aes(x = factor(year))) +
  geom_bar(aes(y = CO2_budget), stat = "identity", fill = "steelblue") +
  geom_text(aes(y = CO2_budget, label = round(CO2_budget, 1)), vjust = -0.5) +
  labs(title = "Annual CO2 Budget by Year",
       x = "Year",
       y = "CO2 Budget (gC/m²/year)") +
  theme_bw()

ggplot(annual_budgets, aes(x = factor(year))) +
  geom_bar(aes(y = CH4_budget), stat = "identity", fill = "darkgreen") +
  geom_text(aes(y = CH4_budget, label = round(CH4_budget, 2)), vjust = -0.5) +
  labs(title = "Annual CH4 Budget by Year",
       x = "Year",
       y = "CH4 Budget (gC/m²/year)") +
  theme_bw()

# 2. Seasonal patterns
# Create seasonal budget plots
ggplot(seasonal_budgets, aes(x = factor(year), y = CO2_budget, fill = season)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Seasonal CO2 Budgets by Year",
       x = "Year",
       y = "CO2 Budget (gC/m²/season)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Environmental relationships
# Create a correlation plot between fluxes and environmental variables
env_flux_daily <- df_avg %>%
  select(FC_F, FCH4_F, TS_3_1_1, SWC_3_1_1, TA, WS, WD, TA_ERA5, TS_ERA5) %>%
  rename(
    CO2_flux = FC_F,
    CH4_flux = FCH4_F,
    Soil_temp = TS_3_1_1,
    Soil_moisture = SWC_3_1_1,
    Air_temp = TA,
    Wind_speed = WS,
    Wind_Direction = WD,
    Air_temp_ERA5 = TA_ERA5,
    Soil_temp_ERA5 = TS_ERA5
  )

cor_matrix <- cor(env_flux_daily, use = "complete.obs")

# Convert correlation matrix to long format for plotting
cor_long <- as.data.frame(as.table(cor_matrix))
names(cor_long) <- c("Var1", "Var2", "Correlation")

ggplot(cor_long, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Matrix of Fluxes and Environmental Variables")
```
#Same as above but using the HH data 
```{r}
# Load necessary packages
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)

# Calculate annual budgets
annual_budgets <- df %>%
  group_by(year = year(date)) %>%
  filter(year != 2023) %>%
  summarize(
    CO2_budget = sum(FC_F * (60*30*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*30*(1/1000000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    mean_air_temp = mean(TA, na.rm = TRUE),
    mean_wind_speed = mean(WS, na.rm = TRUE),
    total_days = n()
  )

# Calculate seasonal budgets
seasonal_budgets <- df %>%
  group_by(year = year(date), season) %>%
  filter(year != 2023) %>%
  summarize(
    CO2_budget = sum(FC_F * (60*30*(1/1000000)*12), na.rm = TRUE),
    CH4_budget = sum(FCH4_F * (60*30*(1/1000000000)*12), na.rm = TRUE),
    GPP_sum = sum(GPP_F * (60*30*(1/1000000)*12), na.rm = TRUE),
    RECO_sum = sum(RECO_F * (60*30*(1/1000000)*12), na.rm = TRUE),
    mean_soil_temp = mean(TS_3_1_1, na.rm = TRUE),
    mean_soil_moisture = mean(SWC_3_1_1, na.rm = TRUE),
    mean_air_temp = mean(TA, na.rm = TRUE),
    mean_wind_speed = mean(WS, na.rm = TRUE),
    days = n()
  )

# Create visualizations

# 1. Annual CO2 and CH4 budgets
ggplot(annual_budgets, aes(x = factor(year))) +
  geom_bar(aes(y = CO2_budget), stat = "identity", fill = "steelblue") +
  geom_text(aes(y = CO2_budget, label = round(CO2_budget, 1)), vjust = -0.5) +
  labs(title = "Annual CO2 Budget by Year",
       x = "Year",
       y = "CO2 Budget (gC/m²/year)") +
  theme_bw()

ggplot(annual_budgets, aes(x = factor(year))) +
  geom_bar(aes(y = CH4_budget), stat = "identity", fill = "darkgreen") +
  geom_text(aes(y = CH4_budget, label = round(CH4_budget, 2)), vjust = -0.5) +
  labs(title = "Annual CH4 Budget by Year",
       x = "Year",
       y = "CH4 Budget (gC/m²/year)") +
  theme_bw()

# 2. Seasonal patterns
# Create seasonal budget plots
ggplot(seasonal_budgets, aes(x = factor(year), y = CO2_budget, fill = season)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Seasonal CO2 Budgets by Year",
       x = "Year",
       y = "CO2 Budget (gC/m²/season)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Environmental relationships
# Create a correlation plot between fluxes and environmental variables
env_flux_daily <- df_avg %>%
  select(FC_F, FCH4_F, TS_3_1_1, SWC_3_1_1, TA, WS) %>%
  rename(
    CO2_flux = FC_F,
    CH4_flux = FCH4_F,
    Soil_temp = TS_3_1_1,
    Soil_moisture = SWC_3_1_1,
    Air_temp = TA,
    Wind_speed = WS
  )

cor_matrix <- cor(env_flux_daily, use = "complete.obs")

# Convert correlation matrix to long format for plotting
cor_long <- as.data.frame(as.table(cor_matrix))
names(cor_long) <- c("Var1", "Var2", "Correlation")

ggplot(cor_long, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Matrix of Fluxes and Environmental Variables")
```

