---
title: "Council_IncMSE_NEE" #script for exploring deconstructed incMSE models for correlated variables 
#Note: optimal OOB determined in BASE data prepping proj, script 1.5 "Optimizing RF ntree OOB"
#Note**: scroll to bottom to see incMSE with full dataset (not train vs test) using cc of all variables being used**
output: html_document
date: "2025-09-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)
library(car) 
library(randomForest)

Sys.setenv(TZ='UTC')
```


#ERA dataset 
####Prepping the era dataset to use to train RF too -- calc vpd and RF for SWC gapfilling, just like in the processing steps 
```{r}
#ERA-5 dataset for RF training - just like in step 2 of the processing code -- uses the same dataset used in the beginning of step 2 
df_era = fread("C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023_era.csv") #using the re-cleaned df from Dani 

#subset out by which has complete air t data 
df_era = df_era[complete.cases(df_era$airt.eramod),]

#create subset of the variables wanted for training RF 
df_era2 = data.frame(df_era$date,df_era$FC.c,df_era$FCH4.c,
                 df_era$airt.eramod, df_era$wd, df_era$cardinal_direction, df_era$rh.eramod,df_era$rad.eramod,
                 df_era$ws.eramod,df_era$tsoil.eramod,#df_era$SWC_1_1_1.c,df_era$SWC_2_1_1.c, 
                 df_era$SWC_3_1_1.c, #SWC 3 = the tussock location 
                 df_era$h.eramod,df_era$le.eramod)

#rename for easier names
names(df_era2) = c('date','nee',"fch4",'tair', 'wd', 'cardDir','rh','rg','ws','tsoil','swc','h','le')

#make card dir as factor - 8 levels 
df_era2$cardDir <- as.factor(df_era2$cardDir)

#calculate VPD from air t and RH
svp = 610.7*10^((7.5*df_era2$tair)/(237.3+df_era2$tair))
df_era2$vpd = ((100 - df_era2$rh)/100)*svp  
```


#RF for SWC - Train RF to fill in SWC (using SWC3, tussocklocation*)
```{r}

#Subset dataset into training and testing dataset 

set.seed(123)#sets the start point of models, good for repeatability
cc = df_era2[complete.cases(df_era2$swc),]#create a gap free data set of the target variable

#use 80% of data set as training set and 20% as test set
sample = sample(c(TRUE, FALSE), nrow(cc), replace=TRUE, prob=c(0.8,0.2))
train  = cc[sample, ]
test   = cc[!sample, ]
#compare to ensure all data sets are representative of total data
hist(cc$swc)
hist(train$swc)
hist(test$swc)

#run random forest to predict missing SWC values 
library(randomForest)

rf.swc = randomForest(formula = swc ~ tair + rh + rg + ws + wd + tsoil + vpd + le + h,data = train,ntree = 800) #optimum ntree determined by OOB error (script in Council Base data prepping, #1.5)

#predict it on the full data set
swc.rf = predict(object = rf.swc,newdata = df_era2)
df_era2$swc.rf = swc.rf 

#time series plot of gap filled vs real
ggplot(data = df_era2)+
  geom_point(aes(date,swc.rf*100,col='RF'))+
  geom_point(aes(date,swc*100,col='Measured'))+
  scale_y_continuous("SWC (%)")


#validation dataset from only test data
val = merge(test,df_era2,by = "date",all.x = T)

ggplot(data = val,aes(swc.rf,swc.x))+theme_bw()+geom_abline(slope = 1,intercept = 0,col='red')+
  geom_point(alpha=0.1)+
  scale_x_continuous(limits = c(0,70),"RF SWC (%)")+
  scale_y_continuous(limits = c(0,70),"Measured SWC (%)")


#summary stats on gap filling
summary(lm(val$swc.x ~ val$swc.rf)) #R2 = 0.98, slope = 1.02

#Fig for council paper supp material 
SWC_RF <- ggplot(data = val,aes(swc.rf,swc.x))+theme_bw()+geom_abline(slope = 1,intercept = 0,col='red')+
  geom_point(alpha=0.1)+
  scale_x_continuous(limits = c(0,70),"RF SWC (%)")+
  scale_y_continuous(limits = c(0,70),"Measured SWC (%)")+
    annotate(geom = 'text',x = 10, y = 70,label=expression(R^2~"= 0.98"),size = 5)+
  annotate(geom = 'text',x = 10,y = 65,label=expression(Slope~"= 1.02"), size = 5)
SWC_RF

# ggsave("C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/SWC_RF_Validation.png", SWC_RF, 
#        width = 10, height = 7, dpi = 600)

#create final gap free soil moisture
df_era2$swc = ifelse(is.na(df_era2$swc),df_era2$swc.rf,df_era2$swc)

```

#SWC-soil temp interaction term 
```{r}
# Create interaction term --> explicit interaction term while keeping original terms 
df_era2$TS_SWC_interact <- df_era2$tsoil * df_era2$swc
```


# VarImp for NEE

#testing for correlated variables 
#Non-parametric tests of correlations - spearman's correlation 

##### Full model with all variables & interaction term 
```{r}
#interaction term in ERA5 dataset, for reference: TS_SWC_interact

#Create training and testing datasets 
cc_nee = df_era2[complete.cases(df_era2$nee),]
cc = subset(cc,cc$nee < 12 & cc$nee > -14) #removing some spread 

#create training dataset - 80% to train, 20% to test
sample.nee = sample(c(TRUE, FALSE), nrow(cc_nee), replace=TRUE, prob=c(0.8,0.2))
train.nee  = cc_nee[sample.nee, ]
test.nee   = cc_nee[!sample.nee, ]


rf.nee_full <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + 
                             vpd + swc + h + le + TS_SWC_interact ,
                             data = train.nee,
                             ntree = 950,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + rg + ws + wd + tsoil + vpd + swc + h + le + TS_SWC_interact, data = df_era2)

vif(vif_model) 
      #      tair              rh              rg              ws              wd           tsoil 
      # 10.876716        4.230284       14.027880        1.241240        1.056183       33.772944 
      #       vpd             swc               h              le TS_SWC_interact 
      # 13.803276        1.842613        6.545965        6.066574       22.582958 

# Extract variable importance
importance_values <- importance(rf.nee_full, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_full, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#                   %IncMSE
# tair             43.83793
# rh               83.60214
# rg               50.46131
# ws               92.66371 *
# wd              113.24795 *
# tsoil            52.33887
# vpd              44.72077
# swc             117.01935 * 
# h                81.75600
# le               62.76682
# TS_SWC_interact  48.44485

```
#full model but using the categorical CarDir instead of wd (degrees)
```{r}
rf.nee_full2 <- randomForest(nee ~ tair + rh + rg + ws + cardDir + tsoil + 
                             vpd + swc + h + le + TS_SWC_interact ,
                             data = train.nee,
                             ntree = 950,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + rg + ws + cardDir + tsoil + vpd + swc + h + le + TS_SWC_interact, data = df_era2)

vif(vif_model) 
#                      GVIF Df GVIF^(1/(2*Df))
# tair            11.115906  1        3.334052
# rh               4.349437  1        2.085530
# rg              14.265122  1        3.776920
# ws               1.326549  1        1.151759
# cardDir          1.554770  7        1.032026
# tsoil           34.096096  1        5.839186
# vpd             13.877429  1        3.725242
# swc              1.913614  1        1.383334
# h                6.921858  1        2.630942
# le               6.156193  1        2.481168
# TS_SWC_interact 22.802175  1        4.775162

# Extract variable importance
importance_values <- importance(rf.nee_full2, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_full2, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#                   %IncMSE
# tair             39.62552
# rh               83.62718
# rg               50.22527
# ws               90.84175 *
# cardDir         120.99655 *
# tsoil            53.93861
# vpd              43.29743
# swc             114.13806 *
# h                77.15994
# le               59.64057
# TS_SWC_interact  48.14305

```


#Boruta variable importance
```{r}
#added wd (degrees)
library(Boruta)

boruta = Boruta(nee ~ tair + rh + rg + ws + wd + tsoil + swc + vpd + le + h + TS_SWC_interact, data = train.nee ,doTrace = 2,maxRuns = 100)
boruta$finalDecision #indicates all above are confirmed important variables via a list 
plot(boruta,las = 2) 

#using carDir instead of wd 
boruta2 = Boruta(nee ~ tair + rh + rg + ws + cardDir + tsoil + swc + vpd + le + h + TS_SWC_interact, data = train.nee ,doTrace = 2,maxRuns = 100)
boruta2$finalDecision #indicates all above are confirmed important variables via a list 
plot(boruta2,las = 2) 
```


#Checking collinearity among continuous predictor variables 
```{r}
library(corrplot)
library(car)
library(dplyr)
# CORRELATION MATRIX - Visual inspection
# Predictor variables
predictors <- df_era2 %>%
  select(tair, rh, rg, ws, wd, tsoil, vpd, swc, h, le, TS_SWC_interact) %>%
  na.omit()  # Remove rows with any missing values

# Calculate correlation matrix
cor_matrix <- cor(predictors, use = "complete.obs")
print("Correlation Matrix:")
print(round(cor_matrix, 3))

# Visualize correlation matrix
corrplot(cor_matrix, method = "color", type = "upper", 
         order = "hclust", tl.cex = 0.8, tl.col = "black",
         addCoef.col = "black", number.cex = 0.7)
title("Collinearity Check: Correlation Matrix")

# Identify highly correlated pairs (|r| > 0.7) - code from claude
high_cor_pairs <- which(abs(cor_matrix) > 0.7 & abs(cor_matrix) < 1, arr.ind = TRUE)
if(nrow(high_cor_pairs) > 0) {
  print("Variable pairs with |correlation| > 0.7:")
  for(i in 1:nrow(high_cor_pairs)) {
    var1 <- rownames(cor_matrix)[high_cor_pairs[i,1]]
    var2 <- colnames(cor_matrix)[high_cor_pairs[i,2]]
    cor_val <- cor_matrix[high_cor_pairs[i,1], high_cor_pairs[i,2]]
    cat(paste(var1, "vs", var2, ": r =", round(cor_val, 3), "\n"))
  }
} else {
  print("No variable pairs with |correlation| > 0.7 found")
}
#Results: most of them have collinearity / correlation 
# strongest correlations: 
#swc - tsoil, TS interact, tair, vpd
#tsoil - TS interact, tair, vpd, le
#TS interact - tair, vpd, le 
#tair - vpd, le, rg, h
#vpd - le, rg, h, rh
#wd - ok
#le - rg, h, rh
#rg - h, rh
#h - rh
#ws - ok 
```


#Now we start removing / swapping out variables that appear to be collinear (see VIF, corr matrix). My approach is to remove the variables with the highest VIF after each model run


#reduced model 1: removing soil temp - soil moisture interaction 
```{r}

rf.nee_1 <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + 
                             vpd + swc + h + le,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + rg + ws + wd + tsoil + vpd + swc + h + le, data = df_era2)

 vif(vif_model)
#      tair        rh        rg        ws        wd     tsoil       vpd       swc         h        le 
# 10.096590  4.111273 13.963473  1.216231  1.050951  7.784172 13.640742  1.820875  6.513567  6.064830 

# Extract variable importance
importance_values <- importance(rf.nee_1, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_1, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tair   40.75993
# rh     80.72256
# rg     46.56342
# ws     94.67850 *
# wd    115.11853 *
# tsoil  73.70873
# vpd    41.20033
# swc   119.08408 *
# h      73.92875
# le     61.35482
```
#Reducing the model by removing the variable with the highest VIF from previous model run 

#reduced model 2: removing soil moist-temp interaction & rg 
```{r}

rf.nee_2 <- randomForest(nee ~ tair + rh + vpd + ws + wd + tsoil + swc + h + le,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + vpd + ws + wd + tsoil + swc + h + le, data = df_era2)

 vif(vif_model) 
 #    tair        rh       vpd        ws        wd     tsoil       swc         h        le 
 # 8.512882  3.960500 13.583562  1.198464  1.050291  6.231644  1.715788  2.844975  4.574293 

# Extract variable importance
importance_values <- importance(rf.nee_2, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_2, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tair   39.40103
# rh     83.33405
# vpd    43.06318
# ws     99.84045
# wd    121.36225 *
# tsoil  77.76151
# swc   137.31217 *
# h     126.56098 *
# le     85.68831
```

#reduced model 3: removing soil moist-temp interaction, rg, & vpd  
```{r}

rf.nee_3 <- randomForest(nee ~ tair + rh + ws + wd + tsoil + swc + h + le,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + ws + wd + tsoil + swc + h + le, data = df_era2)

 vif(vif_model) 
#     tair       rh       ws       wd    tsoil      swc        h       le 
# 5.598534 1.736650 1.122805 1.047441 5.394139 1.550527 2.812684 3.861942 

# Extract variable importance
importance_values <- importance(rf.nee_3, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_3, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tair   53.41029
# rh     85.05921
# ws    103.74377
# wd    109.41255 *
# tsoil 101.23579
# swc   134.00447 *
# h     118.52890 *
# le     86.67041
```


#reduced model 4: removing soil moist-temp interaction, rg, vpd, tsoil   
```{r}

rf.nee_4 <- randomForest(nee ~ tair + rh + ws + wd + swc + h + le,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + ws + wd + swc + h + le, data = df_era2)

 vif(vif_model) 
#     tair       rh       ws       wd      swc        h       le 
# 2.135658 1.736571 1.109577 1.030410 1.366824 2.801886 3.861494 

# Extract variable importance
importance_values <- importance(rf.nee_4, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_4, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#        %IncMSE
# tair 140.96502 *
# rh    97.27419
# ws   112.57898
# wd   109.35117
# swc  158.35607 *
# h    120.91545 *
# le    98.25168
```

#reduced model 5: removing soil moist-temp interaction, rg, vpd, *Swapping out tair and adding tsoil back in   
```{r}

rf.nee_5 <- randomForest(nee ~ tsoil + rh + ws + wd + swc + h + le,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tsoil + rh + ws + wd + swc + h + le, data = df_era2)

 vif(vif_model) 
#    tsoil       rh       ws       wd      swc        h       le 
# 2.057688 1.716520 1.121317 1.040646 1.549723 2.789873 3.606582 

# Extract variable importance
importance_values <- importance(rf.nee_5, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_5, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tsoil 151.03479 *
# rh     89.97427
# ws    103.01713
# wd    119.76066
# swc   146.04809 *
# h     120.44287 *
# le     92.34000

#tsoil has higher incMSE compared to tair when the other is excluded 
#for both versions, le has the highest VIF so removing that next 
```



#reduced model 6: removing soil moist-temp interaction, rg, *tsoil, and le    
```{r}

rf.nee_6 <- randomForest(nee ~ tair + rh + ws + wd + swc + h,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + ws + wd + swc + h, data = df_era2)

 vif(vif_model) 
#     tair       rh       ws       wd      swc        h 
# 1.740642 1.490084 1.062713 1.029823 1.364722 1.809694 

# Extract variable importance
importance_values <- importance(rf.nee_6, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_6, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#        %IncMSE
# tair 177.55333
# rh    95.45722
# ws   150.43097
# wd   128.82545
# swc  188.60975
# h    313.21372


```

#reduced model 7: removing soil moist-temp interaction, vpd, rg, *tair, and le  (swapping tair and tsoil)  
```{r}

rf.nee_7 <- randomForest(nee ~ tsoil + rh + ws + wd + swc + h,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tsoil + rh + ws + wd + swc + h, data = df_era2)

 vif(vif_model) 
#    tsoil       rh       ws       wd      swc        h 
# 1.795630 1.489880 1.075084 1.038298 1.548490 1.640614 

# Extract variable importance
importance_values <- importance(rf.nee_7, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_7, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#        %IncMSE
# tsoil 199.8100 *
# rh    106.4712
# ws    155.6312 *
# wd    128.4710
# swc   145.2717
# h     368.0150 *


```

#reduced model 8: removing VPD, rg, tsoil, tair, and le -- *adding in soil temp-moist interaction
```{r}

rf.nee_8 <- randomForest(nee ~ TS_SWC_interact + rh + ws + wd + swc + h,
                             data = train.nee,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ TS_SWC_interact + rh + ws + wd + swc + h, data = df_era2)

 vif(vif_model) 
# TS_SWC_interact              rh              ws              wd             swc               h 
#        1.718131        1.492326        1.062432        1.034974        1.539461        1.601862 

# Extract variable importance
importance_values <- importance(rf.nee_8, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_8, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#                  %IncMSE
# TS_SWC_interact 175.5376 *
# rh              107.1026
# ws              147.1569
# wd              131.7755
# swc             147.7557 *
# h               357.1714 *


#Notes: 
#*tsoil has higher incMSE compared to interaction term; tair inMSE also slightly higher 

```
# ***********************


#Re-running with cc of ALL variables being testsed, and using the whole dataset (not training vs esting) - this uses the finalized dataset used in analyses 



#complete cases of all predictor variables being used and complete cases of obs nee (not gapfilled)

```{r}
#Complete cases of nee 
cc_nee = df_era2[complete.cases(df_era2$nee),]
cc = subset(cc_nee,cc_nee$nee < 12 & cc_nee$nee > -14) #removes some outliers / spread


#subset out cc
cc_all = cc[complete.cases(cc$tair, cc$wd, cc$cardDir, cc$rh, cc$rg, cc$ws, cc$tsoil, cc$swc, cc$h, cc$le, cc$vpd),]

sum(is.na(cc_all$tair))
```



#SWC-soil temp interaction term 
```{r}
# Create interaction term --> explicit interaction term while keeping original terms 
cc_all$TS_SWC_interact <- cc_all$tsoil * cc_all$swc
```


# VarImp for NEE

#testing for correlated variables 
#Non-parametric tests of correlations - spearman's correlation 

##### Full model with all variables & interaction term, cc of ALL variables 
```{r}

rf.nee_full.cc <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + 
                             vpd + swc + h + le + TS_SWC_interact ,
                             data = cc_all,
                             ntree = 950,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + rg + ws + wd + tsoil + vpd + swc + h + le + TS_SWC_interact, data = cc_all)

vif(vif_model) 
#            tair              rh              rg              ws              wd 
#       10.876306        4.230133       14.025554        1.241197        1.056193 
#           tsoil             vpd             swc               h              le 
#       33.772879       13.803904        1.842530        6.544763        6.066360 
# TS_SWC_interact 
#       22.582830 

# Extract variable importance
importance_values <- importance(rf.nee_full.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_full.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#                   %IncMSE
# tair             44.04506
# rh               87.83592
# rg               50.09184
# ws              106.12816
# wd              141.06816
# tsoil            61.94372
# vpd              50.11790
# swc             127.16900
# h                82.96411
# le               62.92248
# TS_SWC_interact  61.33962

```
#Now we start removing / swapping out variables that appear to be collinear (see VIF, corr matrix). My approach is to remove the variables with the highest VIF after each model run


#reduced model 1: removing soil temp - soil moisture interaction 
```{r}

rf.nee_1.cc <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + 
                             vpd + swc + h + le,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + rg + ws + wd + tsoil + vpd + swc + h + le, data = cc_all)

 vif(vif_model)
#       tair        rh        rg        ws        wd     tsoil       vpd       swc         h 
# 10.096192  4.111179 13.961169  1.216213  1.050961  7.784100 13.641347  1.820802  6.512390 
#        le 
#  6.064610 

# Extract variable importance
importance_values <- importance(rf.nee_1.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_1.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tair   43.47467
# rh     90.13754
# rg     45.32025
# ws    112.47298 *
# wd    131.55486 *
# tsoil  76.82326
# vpd    46.70014
# swc   140.57965 *
# h      86.72934
# le     62.10891
```
#removing variable with highest VIF from previous model 

#reduced model 2: removing soil moist-temp interaction & vpd
```{r}

rf.nee_2.cc <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + swc + h + le,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + rg + ws + wd + tsoil + swc + h + le, data = cc_all)

 vif(vif_model) 
 #     tair        rh        rg        ws        wd     tsoil       swc         h        le 
 # 7.465335  1.971089 13.902665  1.145542  1.047940  6.796080  1.672813  6.420426  5.481915 

# Extract variable importance
importance_values <- importance(rf.nee_2.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_2.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tair   44.05508
# rh     98.98468
# rg     48.72340
# ws    107.87091
# wd    146.52951
# tsoil  90.45768
# swc   142.86412
# h      83.62852
# le     64.32201
```
#reduced model 3: removing soil moist-temp interaction, vpd, & rg
```{r}

rf.nee_3.cc <- randomForest(nee ~ tair + rh + ws + wd + tsoil + swc + h + le,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + ws + wd + tsoil + swc + h + le, data = cc_all)

 vif(vif_model) 
#     tair       rh       ws       wd    tsoil      swc        h       le 
# 5.598571 1.736815 1.122798 1.047448 5.394029 1.550447 2.812292 3.861786 

# Extract variable importance
importance_values <- importance(rf.nee_3.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_3.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tair   56.54131
# rh     91.57067
# ws    118.09106
# wd    123.46611 *
# tsoil 107.01640
# swc   147.59590 *
# h     131.88021 *
# le     83.55430
```
#reduced model 4: removing soil moist-temp interaction, rg, vpd, tsoil   
```{r}

rf.nee_4.cc <- randomForest(nee ~ tair + rh + ws + wd + swc + h + le,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + ws + wd + swc + h + le, data = cc_all)

 vif(vif_model) 
#     tair       rh       ws       wd      swc        h       le 
# 2.135656 1.736737 1.109567 1.030415 1.366761 2.801496 3.861341 

# Extract variable importance
importance_values <- importance(rf.nee_4.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_4.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#        %IncMSE
# tair 142.61403 *
# rh    98.42112
# ws   119.32521
# wd   130.03058 *
# swc  186.98190 *
# h    120.00928
# le   100.43484

```



#reduced model 5: removing soil moist-temp interaction, rg, vpd, *Swapping out tair and adding tsoil back in   
```{r}

rf.nee_5.cc <- randomForest(nee ~ tsoil + rh + ws + wd + swc + h + le,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tsoil + rh + ws + wd + swc + h + le, data = cc_all)

 vif(vif_model) 
#    tsoil       rh       ws       wd      swc        h       le 
# 2.057631 1.716666 1.121311 1.040650 1.549643 2.789477 3.606361 

# Extract variable importance
importance_values <- importance(rf.nee_5.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_5.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#         %IncMSE
# tsoil 155.24056
# rh     95.72478
# ws    117.09667
# wd    126.92049
# swc   160.64890
# h     129.41520
# le     95.42372

#tsoil has higher incMSE compared to tair when the other is excluded 
#for both models, le has the highest VIF so removing that next 

```

#reduced model 6: removing soil moist-temp interaction, rg, *tsoil, and le    
```{r}

rf.nee_6.cc <- randomForest(nee ~ tair + rh + ws + wd + swc + h,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tair + rh + ws + wd + swc + h, data = cc_all)

 vif(vif_model) 
#     tair       rh       ws       wd      swc        h 
# 1.740603 1.490105 1.062742 1.029830 1.364663 1.809669 

# Extract variable importance
importance_values <- importance(rf.nee_6.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_6.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#       %IncMSE
# tair 187.8308 *
# rh   100.1157
# ws   173.3438
# wd   141.8155
# swc  203.8333 *
# h    349.2868 *


```
#reduced model 7: removing soil moist-temp interaction, vpd, rg, *tair, and le  (swapping tair and tsoil)  
```{r}

rf.nee_7.cc <- randomForest(nee ~ tsoil + rh + ws + wd + swc + h,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ tsoil + rh + ws + wd + swc + h, data = cc_all)

 vif(vif_model) 
#    tsoil       rh       ws       wd      swc        h 
# 1.795580 1.489901 1.075118 1.038305 1.548411 1.640588 

# Extract variable importance
importance_values <- importance(rf.nee_7.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_7.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#       %IncMSE
# tsoil 209.8793 *
# rh    111.7399
# ws    169.7729
# wd    147.0569
# swc   172.8441 *
# h     387.5091 *


```

#reduced model 8: removing VPD, rg, tsoil, tair, and le -- *adding in soil temp-moist interaction
```{r}

rf.nee_8.cc <- randomForest(nee ~ TS_SWC_interact + rh + ws + wd + swc + h,
                             data = cc_all,
                             ntree = 850,
                             importance = TRUE)


 vif_model <- lm(nee ~ TS_SWC_interact + rh + ws + wd + swc + h, data = cc_all)

 vif(vif_model) 
# TS_SWC_interact              rh              ws              wd             swc 
#        1.718074        1.492343        1.062468        1.034981        1.539371 
#               h 
#        1.601840 

# Extract variable importance
importance_values <- importance(rf.nee_8.cc, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf.nee_8.cc, type = 1, main = "NEE: Variable Importance (%IncMSE)")

#                  %IncMSE
# TS_SWC_interact 189.7590 *
# rh              110.9547
# ws              176.0384 *
# wd              143.0567
# swc             163.0466
# h               386.0758 *


#Notes: 
#*tsoil has higher incMSE compared to interaction term; tair inMSE also slightly higher 

```
#Testing which model may be better, comparing the reduced models using tair and tsoil 
#### *anova doesn't work on randomforest products 
#### AIC/BIC also don't work for RF models 

```{r}

#using OOB and RMSE to compare the reduced models 6 and 7 

oob_mse_6 <- tail(rf.nee_6.cc$mse, 1)
oob_r2_6  <- tail(rf.nee_6.cc$rsq, 1)

oob_mse_7 <- tail(rf.nee_7.cc$mse, 1)
oob_r2_7  <- tail(rf.nee_7.cc$rsq, 1)

data.frame(model = c("airT", "soilT"),
           OOB_MSE = c(oob_mse_6, oob_mse_7),
           OOB_R2  = c(oob_r2_6,  oob_r2_7))

# # model      OOB MSE               OOB R2
# airT      	2.235797	           0.6386396		
# soilT	      2.173410	           0.6487229	

#tsoil has slightly lower error, and slightly higher R2, indicates it may be the better model 


#looking at model 8 with interaction term 
oob_mse_8 <- tail(rf.nee_8.cc$mse, 1)
oob_r2_8  <- tail(rf.nee_8.cc$rsq, 1)

data.frame(model = c("airT", "soilT", "interaction"),
           OOB_MSE = c(oob_mse_6, oob_mse_7, oob_mse_8),
           OOB_R2  = c(oob_r2_6,  oob_r2_7, oob_r2_8))
#interaction has slightly better performance, very close to tsoil 

# # model      OOB MSE               OOB R2
# airT      	2.235797	           0.6386396		
# soilT	      2.173410	           0.6487229		
# interaction	2.167859	           0.6496200	

```
#comparing lm models of reduced RF models 
```{r}
#lm of rf.nee_7
lm_rf.nee_6<- lm(nee ~ tair + rh + ws + wd + swc + h, data = cc_all)
lm_rf.nee_7 <- lm(nee ~ tsoil + rh + ws + wd + swc + h, data = cc_all)

anova(lm_rf.nee_7,lm_rf.nee_6 )
AIC(lm_rf.nee_7,lm_rf.nee_6) #AIC for 6 = 144611.8, AIC for 7 = 144584.1 so 7 has slightly lower AIC
BIC(lm_rf.nee_7,lm_rf.nee_6) #BIC for 6 = 144679.4, BIC for 7 = 144651.7 so 7 has slightly lower BIC

```

