---
title: "Council_Seasons" #Adding seasons to the datasets - seasons based on findings in code "Council_Seasonal_EC_Fluxes.Rmd" 
output: html_document
date: "2025-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Notes on site: Wind direction primarily blows from NW, except for Dec - Mar, when it primarily blows from SE direction. The site is largely tussock tundra but to the south there is thermokarst 
#FC is in umol/m2/s, FCH4 in nano-mol/m2/s 


```{r, include=FALSE}
rm(list = ls())

library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(pracma)
library(dplyr)


Sys.setenv(TZ='UTC')
```

# Set working directory and then Load data

```{r}
#using the ".2" version to reflect the updated SWC used in RF 

#original half-hourly dataframe 
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

#needs continuous day variable - daily avg, calculated in processing steps
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

# #monthly avg, calculated in processing steps 
# df_monthly_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_monthly_AVG_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


```


```{r}
#quick look at the different temp profiles (all at 15 cm depth, diff locations)
sum(is.na(df$TS_1_1_1)) #60277 NA's in temp by margin pond 
sum(is.na(df$TS_2_1_1)) #48729 NA's in temp  by lichen/berries  
sum(is.na(df$TS_3_1_1)) #50445 NA's in temp by tussock 
sum(is.na(df$TS_4_1_1)) #52796 NA's in temp by foot of EC tower
sum(is.na(df$TS_5_1_1)) #new temp in latest upload with 2023 data, don't have info on what it is yet** 


#SWC also broken up by location
#  SWC_1_1_1 % Soil water content (15cm depth) – margin pond
# SWC_2_1_1 % Soil water content (15cm depth) – lichen/berries
# SWC_3_1_1 % Soil water content (15cm depth) - tussock


```

```{r}
#checking timestamp 
ggplot(data = df,aes(TIMESTAMP_END,FC_F))+
  geom_point()+
  scale_x_datetime(limits = as.POSIXct(c('2021-09-28','2021-10-05')))+
  scale_y_continuous(limits = c(-12,5))
```

#create a useable timestamp - not necessary here as it's already in the correct format from our BASE processing code 
```{r}
#ignore this if df is already in correct format, as this might mess up the formatting here and cause issues with analysis
# df$TIMESTAMP_END = as.character(df$TIMESTAMP_END)
# df$TIMESTAMP_START = as.character(df$TIMESTAMP_START)
# 
# df$TIMESTAMP_END = as.POSIXct(df$TIMESTAMP_END, tz="UTC", format = "%Y%m%d%H%M")
# df$TIMESTAMP_START = as.POSIXct(df$TIMESTAMP_START, tz="UTC", format = "%Y%m%d%H%M")
```

#Add Seasonal delineations to df and df_avg (see older version of code for work on delineations)
```{r}
# add year column if not already present
df <- df %>%
  mutate(year = format(date, "%Y"))

# Create DOY for df (half-hourly data)
df$DOY <- yday(df$date)


# Create seasonal assignments using both year and DOY
df<- df %>%
  mutate(
    season = case_when(
      # 2017 seasons
      year == "2017" & (DOY >= 258 & DOY <= 290) ~ 'Fall Senescence',
      year == "2017" & (DOY >= 131 & DOY <= 257) ~ 'Growing Season',
      year == "2017" & (DOY >= 291 | DOY <= 130) ~ 'Winter',
      
      
      # 2018 seasons
      year == "2018" &  (DOY >= 257 & DOY <= 300) ~ 'Fall Senescence',
      year == "2018" & (DOY >= 152 & DOY <= 256) ~ 'Growing Season',
      year == "2018" & (DOY >= 301 | DOY <= 151) ~ 'Winter',
      
      # 2019 seasons
      year == "2019" & (DOY >= 246 & DOY <= 284) ~ 'Fall Senescence',
      year == "2019" & (DOY >= 145 & DOY <= 245) ~ 'Growing Season',
      year == "2019" & (DOY >= 285 | DOY <= 144) ~ 'Winter',
      
      # 2020 seasons
      year == "2020" & (DOY >= 252 & DOY <= 299) ~ 'Fall Senescence',
      year == "2020" & (DOY >= 136 & DOY <= 251) ~ 'Growing Season',
      year == "2020" & (DOY >= 300 | DOY <= 135) ~ 'Winter',
      
      # 2021 seasons
      year == "2021" & (DOY >= 260 & DOY <= 283) ~ 'Fall Senescence',
      year == "2021" &(DOY >= 142 & DOY <= 259) ~ 'Growing Season',
      year == "2021" &(DOY >= 284 | DOY <= 141) ~ 'Winter',
      
      # 2022 seasons
      year == "2022" & (DOY >= 245 & DOY <= 281) ~ 'Fall Senescence',
      year == "2022" & (DOY >= 144 & DOY <= 244) ~ 'Growing Season',
      year == "2022" & (DOY >= 282 | DOY <= 143) ~ 'Winter',
      
      # not doing one for 2023 ** incomplete year 
      TRUE ~ NA_character_
    )
  )

####################season to daily avg dataset###########################

# add year column if not already present
df_avg <- df_avg %>%
  mutate(year = format(date, "%Y"))


# Create DOY For df_avg (daily data)
df_avg$DOY <- yday(df_avg$date)


# Create seasonal assignments using both year and DOY
df_avg <- df_avg %>%
  mutate(
    season = case_when(
      # 2017 seasons
      year == "2017" & (DOY >= 258 & DOY <= 290) ~ 'Fall Senescence',
      year == "2017" & (DOY >= 131 & DOY <= 257) ~ 'Growing Season',
      year == "2017" & (DOY >= 291 | DOY <= 130) ~ 'Winter',
      
      
      # 2018 seasons
      year == "2018" &  (DOY >= 257 & DOY <= 300) ~ 'Fall Senescence',
      year == "2018" & (DOY >= 152 & DOY <= 256) ~ 'Growing Season',
      year == "2018" & (DOY >= 301 | DOY <= 151) ~ 'Winter',
      
      # 2019 seasons
      year == "2019" & (DOY >= 246 & DOY <= 284) ~ 'Fall Senescence',
      year == "2019" & (DOY >= 145 & DOY <= 245) ~ 'Growing Season',
      year == "2019" & (DOY >= 285 | DOY <= 144) ~ 'Winter',
      
      # 2020 seasons
      year == "2020" & (DOY >= 252 & DOY <= 299) ~ 'Fall Senescence',
      year == "2020" & (DOY >= 136 & DOY <= 251) ~ 'Growing Season',
      year == "2020" & (DOY >= 300 | DOY <= 135) ~ 'Winter',
      
      # 2021 seasons
      year == "2021" & (DOY >= 260 & DOY <= 283) ~ 'Fall Senescence',
      year == "2021" &(DOY >= 142 & DOY <= 259) ~ 'Growing Season',
      year == "2021" &(DOY >= 284 | DOY <= 141) ~ 'Winter',
      
      # 2022 seasons
      year == "2022" & (DOY >= 245 & DOY <= 281) ~ 'Fall Senescence',
      year == "2022" & (DOY >= 144 & DOY <= 244) ~ 'Growing Season',
      year == "2022" & (DOY >= 282 | DOY <= 143) ~ 'Winter',
      
      # not doing one for 2023 ** incomplete year 
      TRUE ~ NA_character_
    )
  )

#Save dataset with seasons added 

write.csv(df, "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023seasons_for analysis.2.csv", row.names = FALSE)

write.csv(df_avg, "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023seasons_for analysis.2.csv", row.names = FALSE)


```